---
title: "Prevalence, Risk Factors and Renoprotective Treatment of Chronic Kidney Disease Among Internal Medicine Outpatients at a Central European Tertiary Care Centre: A Cross-Sectional Study"
subtitle: "Statistical report"
format: 
  html:  
    embed-resources: true
    keep-md: false
    toc: true
    toc-expand: 3
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 4
    code-fold: true
    code-tools: true
    code-summary: "Open code"
    grid: 
      body-width: 1000px
      sidebar-width: 600px
      margin-width: 100px
editor: source
theme: darkly
highlight-style: github-dark
fontsize: 12 px
code-block-bg: "#2D2D31"
bibliography: ../references.bib
fontcolor: "#D1D1D1"
monobackgroundcolor: "#3B3B3B"
---

------------------------------------------------------------------------


```{r setup, echo=FALSE}
path <- '/home/ticf/secured_data/GitRepo/ticf/736_ZAHI_CKD'
knitr::opts_chunk$set(collapse = TRUE)
knitr::opts_knit$set(root.dir = path)
```

```{css, echo=FALSE}

.sourceCode {
  font-size: 0.86em;
  border: 1px solid #ffffff; 
  padding: 9px; 
}

code{
    color: #EEE8AA; 
    background: #96464D
}

.equation {
  text-align: center;
  display: block;
  margin: auto;
}

body {
  text-align: justify;
}
```

{{< include ../README.md >}}

# Initiation

Upload initiation file

```{r}
setwd(path)
source('r/736_initiation.R')
```

## Function to get age-specific CKD cut-off

According to [Delanaye et al.](https://pmc.ncbi.nlm.nih.gov/articles/PMC4810758/)

```{r}
trans_egfr_cut <- function(age, sex){
  if (sex == 'male'){
    cut_point = 86.4* (0.993^age) 
  } else if (sex == 'female'){
    cut_point = 88.2 * (0.993^age) 
  } else {
    cut_point = "Sex is not correctly specified. Choose any of 'male' or 'female'"
  }
  return(cut_point)
}

trans_egfr_cut(20, 'male')
trans_egfr_cut(20, 'female')

trans_egfr_cut(40, 'male')
trans_egfr_cut(40, 'female')

trans_egfr_cut(60, 'male')
trans_egfr_cut(60, 'female')

trans_egfr_cut(80, 'male')
trans_egfr_cut(80, 'female')

trans_egfr_cut(80, 'nevim')
```

## Further data wrangling

### `data_ckd`


`BMI_over30` = Obesity indicator (BMI > 30 kg/m²)   

`sex_female` = Binary indicator for female sex   

`HF_or_AF` = Combined indicator of heart failure or atrial fibrillation   

`eGFR_AgeSpecCut` = Age- and sex-specific eGFR thresholds for CKD definition   

`eGFR_min` = eGFR in mL/min/1.73m² unit   

`GFR_text_AgeSpecCut` = Categorical eGFR status (normal/altered) based on age-calibrated thresholds   

`identified_CKD_by_lab_AgeSpecCut` = CKD status based on laboratory criteria using age-specific thresholds   


```{r}
data_ckd <- data_ckd %>%
  dplyr::mutate(
    BMI_over30 = if_else(BMI_last > 30, 1, 0),
    sex_female = if_else(sex == "F", 1, 0),
    HF_or_AF = if_else(
      (FU_heart_failure == 1 | FU_atrial_fibrillation == 1), 1, 0
    ),
    eGFR_AgeSpecCut = if_else(
      sex == "M", trans_egfr_cut(age, "male"), trans_egfr_cut(age, "female")
    ),
    eGFR_min = eGFR_value * 60
  ) %>%
  dplyr::mutate(
    GFR_text_AgeSpecCut = if_else(
      eGFR_min > eGFR_AgeSpecCut, "normal", "altered"
    )
  ) %>%
  dplyr::mutate(
    identified_CKD_by_lab_AgeSpecCut = if_else(
      (GFR_text_AgeSpecCut == "altered") | (ACR_text == "altered"), 1, 0
    )
  )
```

### `data_survival`

`sex_female` = Binary indicator for female sex   

`HF_or_AF` = Combined indicator of heart failure or atrial fibrillation   

`FU_atherosclerosis` = Combined indicator of atherosclerotic disease (ischemic heart disease, other atherosclerosis, peripheral artery disease, or stroke)   


```{r}
data_survival <- data_survival %>%
  dplyr::mutate(
    sex_female = if_else(Sex == "F", 1, 0),
    HF_or_AF = pmax(FU_heart_failure, FU_atrial_fibrillation),
    FU_atherosclerosis = pmax(
      ischemic_heart_disease,
      FU_atherosclerosis_PAD,
      FU_stroke
    )
  )
```

### `data_ckd_complete`

Only these rows with BMI data (30 cases are thus excluded).
Consent as factor

```{r}
data_ckd_complete <- data_ckd %>%
  dplyr::filter(
    !is.na(BMI_over30)
  ) %>%
  dplyr::mutate(
    consent = factor(consent)
  )
```


## Explore `data_ckd` data

```{r}
skimr::skim(data_ckd)
summary(data_ckd)
```


### Comapre cohorts

`consent == 1` indicates *screening cohort* whereas `consent == 0` indicates unscreened patient

```{r}
data_ckd %>%
  dplyr::mutate(consent = if_else(consent == 1, 'consent', 'no_consent')) %>% 
  dplyr::select(
    consent:sex,
    uACR_value:identified_CKD_by_lab_AgeSpecCut
  ) %>%
  gtsummary::tbl_summary(by = consent) %>% 
  add_p()
```

### Number of people based on CKD

#### Select data

```{r}
data_ckd_sum <- data_ckd %>% 
  dplyr::mutate(
    ckd_status = case_when(
      (identified_CKD_by_lab == 1) & (identified_CKD_by_lab_AgeSpecCut == 1) ~ 
        'CKD_both',
      (identified_CKD_by_lab == 1) & (identified_CKD_by_lab_AgeSpecCut == 0) ~ 
        'CKD_only_original',
      (identified_CKD_by_lab == 0) & (identified_CKD_by_lab_AgeSpecCut == 1) ~ 
        'CKD_only_age_calibrated',
      (identified_CKD_by_lab == 0) & (identified_CKD_by_lab_AgeSpecCut == 0) ~ 
        'NO CKD',
    )
  ) %>% 
  dplyr::select(
    consent:uACR_value, eGFR_value,
    identified_CKD_by_lab:HF_or_AF, eGFR_AgeSpecCut, ckd_status,
    ACR_text, eGFR_text, GFR_text_AgeSpecCut
    )
```

#### Data summary - both cohorts

```{r}
data_ckd_sum %>% 
  gtsummary::tbl_summary(by = ckd_status)
```

#### Data summary - screening cohort

```{r}
data_ckd_sum %>% 
  dplyr::filter(consent == 1) %>% 
  dplyr::select(-consent) %>% 
  gtsummary::tbl_summary(by = ckd_status)
```


# Correlations

## Overall correlations

```{r, fig.height=7, fig.width=8}
M <- data_ckd %>%
  dplyr::select(
    where(is.numeric)
  ) %>%
  dplyr::select(
    age:FU_finerenone, FU_stroke:BMI_over30, 
    CKD_record:identified_CKD_by_lab, HF_or_AF
  ) %>%
  cor(method = "spearman", use = "pairwise.complete.obs")

kableExtra::kable(M)

par(mfrow = c(1, 1))

corrplot::corrplot(
  M,
  method = "square",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  tl.col = "black",
  tl.cex = 0.8,
  tl.srt = 40,
  diag = TRUE,
  order = "hclust",
  hclust.method = "average",
)
```

## Correlations between canditate predictors in multivariable models

```{r, fig.height=8, fig.width=10}

M <- data_ckd %>%
  dplyr::select(
    where(is.numeric)
    ) %>%
  dplyr::select(
    age:eGFR_value,
    transplant, BMI_last, BMI_over30,  FU_stroke:FU_charlson_index, FU_stroke, 
    HF_or_AF
  ) %>% 
  cor(method = "spearman", use = "pairwise.complete.obs")


corrplot::corrplot(
  M,
  method = "number",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  tl.col = "black",
  tl.cex = 0.8,
  tl.srt = 40,
  diag = TRUE,
  order = "hclust",        
  hclust.method = "average",
)
```

## UpSet plot for comorbidities

Only these combinations of comorbodities present in at least 30 people are shown

```{r, fig.height=5.5, fig.width=8}

data_upset <- data_ckd %>%
  dplyr::mutate(
    age_60plus = if_else(age >= 60, 1, 0),
    BMI_over30 = if_else(is.na(BMI_over30), 0, BMI_over30),
    CKD = if_else(identified_CKD_by_lab == 1, "Yes", "No")
  ) %>%
  dplyr::select(
    transplant,
    FU_DM1, FU_DM2,
    FU_steatosis, FU_hypertension, HF_or_AF,
    FU_atherosclerosis, CKD
  ) %>%
  data.frame()

names(data_upset)
namely <- c(
  "Non-renal TX", "T1DM", "T2DM", "Liver steatosis",
  "A. hypertension", "HF or AF",
  "ASCVD", "CKD"
)

names(data_upset) <- namely

sets <- names(data_upset)[names(data_upset) != "CKD"]

my_themes <- ComplexUpset::upset_modify_themes(list(
  overall_sizes = theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  ),
  intersections_matrix = theme(
    axis.text.y = element_text(size = 11)
  ),
  `Intersection size` = theme(
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 13)
  )
))

p <- ComplexUpset::upset(
  data_upset,
  intersect = sets,
  set_sizes = ComplexUpset::upset_set_size() +
   scale_x_continuous(labels = NULL),
  
  base_annotations = list(
    "Intersection size" = ComplexUpset::intersection_size(
      text = list(size = 0)
    )
  ),
  
  annotations = list(
    "CKD proportion" = (
      ggplot(mapping = aes(fill = CKD)) +
        geom_bar(position = "fill") +
        scale_y_continuous(
          labels = percent_format(accuracy = 1),
          breaks = c(0, 1),
        ) +
        scale_fill_manual(
          values = c(
            "Yes" = "red",
            "No" = "grey80"
          )
        ) +
        labs(y = "CKD", x = NULL, fill = "CKD") +
        theme(
          axis.text.y = element_text(size = 12),
          axis.title.y = element_text(size = 13)
        )
    )
  ),
  
  width_ratio = 0.2,
  height_ratio = 0.6,
  min_size = 30,
  themes = my_themes
)

plotac <- "sup_figure_upset_comorbidities"
path <- "gitignore/figures"

fig <- p + patchwork::plot_layout(
  heights = c(
    0.5,
    1,
    0.7
  )
)

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 8,
    height = 5.5
  )
}
```

# Precision-based sample size calculation

Select pre-specified data

```{r}
data_power <- data_ckd %>%
  dplyr::filter(
    date_sampling < ymd("2025-03-01"),
    consent == 1
  ) %>%
  dplyr::select(identified_CKD_by_lab)

summary(data_power)
dim(data_power)
```

Estimated proportion and CI

```{r}
ci <- inv_logit(
  confint(
    glm(
      identified_CKD_by_lab ~ +1,
      data = data_power,
      family = binomial(link = "logit")
    )
  )
)

ci

prop <- mean(data_power$identified_CKD_by_lab)
prop
```

## 90% probability that 95% CI width ≤ 5%

```{r}
plotac <- "suppl_fig_1"
path <- "gitignore/figures"

sim_prec <- function(size, probs, reps = 2000, ci_width = 0.05) {
  x <- rbinom(n = reps, size = size, prob = probs)

  ciw <- vapply(
    x,
    function(xx) {
      ci <- binom.test(xx, size)$conf.int
      ci[2] - ci[1]
    },
    numeric(1)
  )

  mean(ciw <= ci_width)
}

sizes <- seq(1000, 2000, by = 25)

res <- mclapply(
  sizes,
  function(n) {
    data.frame(
      n = n,
      precision_probability = sim_prec(
        n,
        probs = prop,
        reps = 2000
      )
    )
  },
  mc.cores = detectCores() - 1
) %>%
  dplyr::bind_rows()

kableExtra::kable(res)

fig <- ggplot(
  res,
  aes(
    x = n,
    y = precision_probability
  )
) +
  geom_line() +
  geom_point() +
  geom_hline(
    yintercept = 0.9,
    linetype = 2
  ) +
  scale_x_continuous(
    breaks = seq(1000, 2000, by = 100)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = "Sample size",
    y = "Probability of achieving target precision"
  ) +
  theme_minimal()

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 8,
    height = 5.5
  )
}

```

$\Rightarrow$ According to the defined criterion, a total sample size of 1,425 participants was sufficient: it was the smallest evaluated sample size at which more than 90% of the 2000 simulated datasets produced a 95% confidence interval narrower than 5 percentage points. Larger sample sizes (e.g., 1,450) provided an even higher probability that the interval width would be below this threshold, but were not required to meet the predefined target.

# Models

## Identified_CKD_by_lab

```{r}
out_mean <- logit(mean(data_ckd$identified_CKD_by_lab))
sd(log2(data_ckd$FU_charlson_index+1))
```

### Unadjsuted effects

#### Define predictors

```{r}
preds <- c(
    'age_30y', 'age_60plus', 'sex_female', 'BMI10_last', 'BMI_over30', 'transplant', 
    'FU_DM1', 'FU_DM2', 'FU_DM', 
    'FU_atherosclerosis', 'FU_stroke', 'FU_PAD', 'ischemic_heart_disease', 
    'FU_hypertension', 'FU_atrial_fibrillation','FU_heart_failure', 'HF_or_AF',
   'log2_FU_charlson_index', 'charlson_index_5', 'FU_steatosis'
)

```

#### Fit models

```{r}
kableExtra::kable(
  cbind(
    unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age / 30),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          sex_female = if_else(sex == "F", 1, 0),
          BMI10_last = (BMI_last / 10)
        ) %>% dplyr::filter(
          consent == 0
        ),
      outcome = "identified_CKD_by_lab",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>% rename(
      `OR [95% CI]` = `OR`
    ),
    unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age / 30),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          sex_female = if_else(sex == "F", 1, 0),
          BMI10_last = (BMI_last / 10)
        ) %>% dplyr::filter(
          consent == 1
        ),
      outcome = "identified_CKD_by_lab",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>%
      dplyr::select(OR, N) %>%
      rename(
        `OR [95% CI]` = `OR`
      ),
    unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age / 30),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          BMI10_last = (BMI_last / 10)
        ),
      outcome = "identified_CKD_by_lab",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>%
      dplyr::select(OR, N) %>%
      rename(
        `OR [95% CI]` = `OR`
      )
  ),
  align = c("r", "c", "c", "c", "c", "c", "c")
) %>%
  add_header_above(
    c(" " = 1, "unscreened" = 2, "screening" = 2, "enriched" = 2)
  ) %>%
  column_spec(1, width = "10em") %>%
  column_spec(c(2, 4, 6), width = "8em") %>%
  column_spec(c(3, 5, 7), width = "5em")

```

#### Visualisation

```{r, fig.height=6.302, fig.width=6.02}

 unadj_eff_con0 <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 0
        ),
      outcome = "identified_CKD_by_lab",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )

 unadj_eff_con1 <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 1
        ),
      outcome = "identified_CKD_by_lab",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )
 
  unadj_eff_overall <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ),
      outcome = "identified_CKD_by_lab",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )
 
df_plot <- bind_rows(
  unadj_eff_con1 %>% mutate(consent = "screening"),
  unadj_eff_overall %>% mutate(consent = "overall")
) %>%
  mutate(
    predictor = recode(
      predictor,
      "age_30y" = "Age (per 30y)",
      "age_60plus" = "Age ≥60y",
      "sex_female" = "Female sex",
      "BMI10_last" = "BMI (per 10 kg/m²)",
      "BMI_over30" = "BMI ≥30 kg/m²",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_DM" = "Any diabetes",
      "FU_atherosclerosis" = "ASCVD",
      "FU_stroke" = "Stroke",
      "FU_PAD" = "Peripheral artery disease",
      "ischemic_heart_disease" = "Ischemic heart disease",
      "FU_hypertension" = "Arterial hypertension",
      "FU_atrial_fibrillation" = "Atrial fibrilation",
      "FU_heart_failure" = "Heart failure",
      "HF_or_AF" = "Heart failure or atrial fibrilation",
      "log2_FU_charlson_index" = "log₂ (Charlson index +1)",
      "charlson_index_5" = "Charlson index ≥5",
      "FU_steatosis" = "Liver steatosis"
    ),
    predictor = factor(predictor, levels = rev(unique(predictor))),
    consent = factor(
  consent,
  levels = c("screening", "overall", "enriched"),
  labels = c("Screening cohort", "Enriched cohort", "No consent cohort")
)
  )
    
  
ggplot(df_plot, aes(x = OR, y = predictor, color = consent, group = consent)) +
  geom_point(
    position = position_dodge(width = 0.6),
    size = 2
    ) +
  geom_errorbarh(
    aes(xmin = CI_L, xmax = CI_U),
    position = position_dodge(width = 0.6),
    height = 0.3
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray40") +
  scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  scale_color_manual(
    values = c(
      "Enriched cohort" = "skyblue3",
      "Screening cohort" = "#d95f02",
      "No consent cohort" = "#1b9e77"
    ), 
    breaks = c(
      "Enriched cohort", "Screening cohort", "No consent cohort")
  ) +
  labs(
    x = "Unadjusted odds ratio (log2 scale) for confirmed CKD",
    y = NULL,
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    axis.text.y = element_text(size = 11)
  )

```

### Simple prevalence model

#### Prior

```{r}
prior_intercept <- prior_string(
      paste0("student_t(3, ", out_mean, ", 100)"), 
      class = "Intercept"
    )

```

#### Fit simple mdoel

```{r}
model_simple <- run(
  expr = brm(
    identified_CKD_by_lab ~ +1,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    prior = prior_intercept,
    cores = 1, chains = 2,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_simple",
  reuse = TRUE
)

prior_summary(model_simple)
summary(model_simple, robust = TRUE)

kableExtra::kable(
  inv_logit(fixef(model_simple, robust = TRUE)[, c(1, 3,4)])
)
```

### Consent-specific

```{r}


model_consent_spec <- run(
  expr = brm(
    identified_CKD_by_lab ~ consent,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = prior_intercept,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_consent_spec",
  reuse = TRUE
)


prior_summary(model_consent_spec)

summary(model_consent_spec, robust = TRUE)

posterior <- as_draws_df(model_consent_spec) %>% 
  mutate(
    consent_prev = inv_logit(b_Intercept + b_consent),
    nonconsent_prev = inv_logit(b_Intercept)
    )

tr <- posterior_epred(model_consent_spec)
tr2 <- rowMeans(tr)

kableExtra::kable(
  data.frame(
    rbind(
      inv_logit(fixef(model_simple, robust = TRUE)[, c(1, 3, 4)]),
      quantile(tr2, probs = c(0.5, 0.025, 0.975)),
      quantile(posterior$nonconsent_prev, probs = c(0.5, 0.025, 0.975)),
      quantile(posterior$consent_prev, probs = c(0.5, 0.025, 0.975))
    )
  ) %>%
    dplyr::mutate(
      Estimand = c(
        "Marginal prevalence - simple model",
        "Marginal prevalence - consent-adj. model",
        "Prevalence - consent=0",
        "Prevalence - consent=1"
      )
    ) %>% dplyr::select(
      Estimand, Estimate, Q2.5, Q97.5
    )
)

quantile(posterior$consent_prev, probs = c(0.5, 0.025, 0.975))
```

### Non-linear time model

#### Fit model

```{r}
model_time_nl <- run(
  expr = brm(
    identified_CKD_by_lab ~
      consent +
      s(I( (age-62) / 30), by = consent, k = 5),
    family = bernoulli(link = "logit"),
    data = data_ckd %>% dplyr::mutate(consent = factor(consent)),
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = prior_intercept,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_time_nl",
  reuse = TRUE
)


prior_summary(model_time_nl)
summary(model_time_nl, robust = TRUE)

```

#### Get posteriors

```{r}
quantile(data_ckd$age, probs = c(0, 0.01, 0.05, 0.25, 0.5, 0.75, 0.95, 0.99))

age_seq <- c(
  seq(min(data_ckd$age),
    max(data_ckd$age),
    length.out = 101
  ),
  c(seq(25, 80, by = 5))
)

consent_prop <- mean(data_ckd$consent)

new0 <- data.frame(
  age = age_seq,
  consent = factor(0, levels = c(0, 1))
)

new1 <- data.frame(
  age = age_seq,
  consent = factor(1, levels = c(0, 1))
)

p0 <- posterior_epred(model_time_nl, newdata = new0)
p1 <- posterior_epred(model_time_nl, newdata = new1)
p_overall <- (1 - consent_prop) * p0 + consent_prop * p1



summarize_epred <- function(mat, cohort) {
  posterior_summary(mat, robust = TRUE) %>%
    as.data.frame() %>%
    dplyr::mutate(
      cohort = cohort,
      age = age_seq
    )
}

prediction_NL <- dplyr::bind_rows(
  summarize_epred(p1, "screening"),
  summarize_epred(p0, "unscreened"),
  summarize_epred(p_overall, "enriched")
  ) %>%
  dplyr::mutate(age_calibrated = "No calibration")


kableExtra::kable(
  prediction_NL %>%
    dplyr::filter(age %in% seq(25, 80, by = 5)) %>% 
    dplyr::select(-age_calibrated)
)

```

#### Visualisation

```{r}
prediction_NL %>% 
  dplyr::filter(cohort != 'unscreened') %>% 
  ggplot(
    aes(
      x = age,
      y = Estimate,
      color = cohort,
      fill = cohort
    )
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = Q2.5,
      ymax = Q97.5
    ),
    alpha = 0.4,
    color = NA
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)
  ) +
  labs(
    x = "Age (years)",
    y = "Estimated probability of CKD"
  ) +
  scale_color_manual(
    values = c(
        "unscreened" = "#1b9e77",
        "screening" = "#d95f02",
        "enriched" = "skyblue3"
    ),
    name = "Cohort"
  ) +
  scale_fill_manual(
     values = c(
        "unscreened" = "#1b9e77",
        "screening" = "#d95f02",
        "enriched" = "skyblue3"
    ),
    name = "Cohort"
  ) +
  facet_grid(rows = vars(cohort)) +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    strip.text.x = element_text(size = 12),
    legend.position = "none"
  )
```

### Risk factors model

#### Can age be lienarized?

```{r}
m_tr <- mgcv::gam(
  identified_CKD_by_lab ~ 
      s(age,  k = 5) +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent,
  data = data_ckd,
  family = binomial(link = 'logit')
  )

summary(m_tr)
plot(m_tr)
```

#### Set prior and fit

```{r}
mean(log2(data_ckd$FU_charlson_index + 1))
2*sd(log2(data_ckd$FU_charlson_index + 1))

sd((log2(data_ckd$FU_charlson_index + 1) - 1.64)/1.65)
mean((log2(data_ckd$FU_charlson_index + 1) - 1.64)/1.65)

priors <- c(
  prior(student_t(3, 0, 100), class = "Intercept"),

  prior(normal(0, 5), class = "b", coef = "IageM62D30"),
  prior(normal(0, 5), class = "b", coef = "transplant"),
  prior(normal(0, 5), class = "b", coef = "sex_female"),
  prior(normal(0, 5), class = "b", coef = "FU_DM1"),
  prior(normal(0, 5), class = "b", coef = "FU_DM2"),
  prior(normal(0, 5), class = "b", coef = "FU_atherosclerosis"),
  prior(normal(0, 5), class = "b", coef = "FU_steatosis"),
  prior(normal(0, 5), class = "b", coef = "FU_hypertension"),
  prior(normal(0, 5), class = "b", coef = "HF_or_AF"),
  prior(normal(0, 5), class = "b", coef = "consent"),
  
  prior(normal(0, 1), class = "b", coef = "IageM62D30:consent"),
  prior(normal(0, 1), class = "b", coef = "consent:transplant"),
  prior(normal(0, 1), class = "b", coef = "consent:sex_female"),
  prior(normal(0, 1), class = "b", coef = "consent:FU_DM1"),
  prior(normal(0, 1), class = "b", coef = "consent:FU_DM2"),
  prior(normal(0, 1), class = "b", coef = "consent:FU_atherosclerosis"),
  prior(normal(0, 1), class = "b", coef = "consent:FU_steatosis"),
  prior(normal(0, 1), class = "b", coef = "consent:FU_hypertension"),
  prior(normal(0, 1), class = "b", coef = "consent:HF_or_AF")
)

model_complex_lin <- run(
  expr = brm(
    identified_CKD_by_lab ~ 
      I( (age-62) / 30) * consent +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      sex_female * consent + 
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent,
    data = data_ckd,
    family = bernoulli(link = "logit"),
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = priors,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_complex_lin",
  reuse = TRUE
)

prior_summary(model_complex_lin)
summary(model_complex_lin, robust = TRUE)

exp(fixef(model_complex_lin, robust = TRUE))[,c(1, 3, 4)]
```

#### Extract posterior


```{r}
consent_prop <- mean(
  data_ckd$consent
)

age_scale <- 25 / 30

post <- posterior::as_draws_df(
  model_complex_lin
)

b_all <- post %>%
  dplyr::select(
    dplyr::starts_with("b_")
  )

base <- b_all %>%
  dplyr::transmute(
    Intercept = b_Intercept,
    age_25y = b_IageM62D30 * age_scale,
    sex_female = b_sex_female,
    transplant = b_transplant,
    FU_DM1 = b_FU_DM1,
    FU_DM2 = b_FU_DM2,
    FU_steatosis = b_FU_steatosis,
    FU_hypertension = b_FU_hypertension,
    FU_atherosclerosis = b_FU_atherosclerosis,
    HF_or_AF = b_HF_or_AF
  )

eff_consent0 <- base

eff_consent1 <- base %>%
  dplyr::mutate(
    Intercept = Intercept + b_all$b_consent,
    age_25y = age_25y + b_all$`b_IageM62D30:consent` * age_scale,
    transplant = transplant + b_all$`b_consent:transplant`,
    FU_DM1 = FU_DM1 + b_all$`b_consent:FU_DM1`,
    FU_DM2 = FU_DM2 + b_all$`b_consent:FU_DM2`,
    FU_steatosis = FU_steatosis + b_all$`b_consent:FU_steatosis`,
    FU_hypertension = FU_hypertension + b_all$`b_consent:FU_hypertension`,
    FU_atherosclerosis = FU_atherosclerosis + b_all$`b_consent:FU_atherosclerosis`,
    HF_or_AF = HF_or_AF + b_all$`b_consent:HF_or_AF`,
    sex_female = sex_female + b_all$`b_consent:sex_female`
  )

eff_consent_mean <- base %>%
  dplyr::mutate(
    Intercept = Intercept + b_all$b_consent * consent_prop,
    age_25y = age_25y + b_all$`b_IageM62D30:consent` * consent_prop * age_scale,
    transplant = transplant + b_all$`b_consent:transplant` * consent_prop,
    FU_DM1 = FU_DM1 + b_all$`b_consent:FU_DM1` * consent_prop,
    FU_DM2 = FU_DM2 + b_all$`b_consent:FU_DM2` * consent_prop,
    FU_steatosis = FU_steatosis + b_all$`b_consent:FU_steatosis` * consent_prop,
    FU_hypertension = FU_hypertension + b_all$`b_consent:FU_hypertension` * consent_prop,
    FU_atherosclerosis = FU_atherosclerosis + b_all$`b_consent:FU_atherosclerosis` * consent_prop,
    HF_or_AF = HF_or_AF + b_all$`b_consent:HF_or_AF` * consent_prop,
    sex_female = sex_female + b_all$`b_consent:sex_female` * consent_prop
  )

calc_marg_log_or <- function(model, data, var_name, val1, val0) {
  
  d1 <- data
  
  d1[[var_name]] <- val1
  
  p1 <- brms::posterior_epred(
    model,
    newdata = d1
  )
  
  p1_marg <- rowMeans(
    p1
  )
  
  d0 <- data
  
  d0[[var_name]] <- val0
  
  p0 <- brms::posterior_epred(
    model,
    newdata = d0
  )
  
  p0_marg <- rowMeans(
    p0
  )
  
  odds1 <- p1_marg / (1 - p1_marg)
  
  odds0 <- p0_marg / (1 - p0_marg)
  
  marg_or <- odds1 / odds0
  
  return(log(marg_or))
}

data_dm_ref <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 0
  )

p0_dm <- brms::posterior_epred(
  model_complex_lin,
  newdata = data_dm_ref
) %>%
  rowMeans()

odds0_dm <- p0_dm / (1 - p0_dm)

data_dm1 <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 1,
    FU_DM2 = 0
  )

p1_dm1 <- brms::posterior_epred(
  model_complex_lin,
  newdata = data_dm1
) %>%
  rowMeans()

odds1_dm1 <- p1_dm1 / (1 - p1_dm1)

data_dm2 <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 1
  )

p1_dm2 <- brms::posterior_epred(
  model_complex_lin,
  newdata = data_dm2
) %>%
  rowMeans()

odds1_dm2 <- p1_dm2 / (1 - p1_dm2)

log_or_dm1_marg <- log(odds1_dm1 / odds0_dm)

log_or_dm2_marg <- log(odds1_dm2 / odds0_dm)

eff_marginal <- dplyr::tibble(
  age_25y = calc_marg_log_or(
    model_complex_lin,
    data_ckd,
    "age",
    75,
    50
  ),
  sex_female = calc_marg_log_or(
    model_complex_lin,
    data_ckd,
    "sex_female",
    1,
    0
  ),
  transplant = calc_marg_log_or(
    model_complex_lin,
    data_ckd,
    "transplant",
    1,
    0
  ),
  FU_DM1 = log_or_dm1_marg,
  FU_DM2 = log_or_dm2_marg,
  FU_steatosis = calc_marg_log_or(
    model_complex_lin,
    data_ckd,
    "FU_steatosis",
    1,
    0
  ),
  FU_hypertension = calc_marg_log_or(
    model_complex_lin,
    data_ckd,
    "FU_hypertension",
    1,
    0
  ),
  FU_atherosclerosis = calc_marg_log_or(
    model_complex_lin,
    data_ckd,
    "FU_atherosclerosis",
    1,
    0
  ),
  HF_or_AF = calc_marg_log_or(
    model_complex_lin,
    data_ckd,
    "HF_or_AF",
    1,
    0
  )
)

summ <- dplyr::bind_rows(
  posterior::summarise_draws(
    eff_consent0,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "unscreened"
    ),
  posterior::summarise_draws(
    eff_consent1,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "screening"
    ),
  posterior::summarise_draws(
    eff_consent_mean,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_std_cond"
    ),
  posterior::summarise_draws(
    eff_marginal,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_marginal"
    )
) %>%
  dplyr::rename(
    predictor = variable,
    CI_L = `2.5%`,
    CI_U = `97.5%`
  ) %>%
  dplyr::mutate(
    OR = exp(
      median
    ),
    CI_L = exp(
      CI_L
    ),
    CI_U = exp(
      CI_U
    )
  )

data_forest <- summ %>%
  dplyr::filter(
    predictor != "Intercept"
  ) %>%
  dplyr::mutate(
    predictor = factor(
      predictor,
      levels = rev(
        unique(
          predictor
        )
      )
    )
  )

tab_ready <- summ %>%
  dplyr::filter(
    predictor != "Intercept"
  ) %>%
  dplyr::mutate(
    label = sprintf(
      "%.2f [%.2f, %.2f]",
      OR,
      CI_L,
      CI_U
    )
  ) %>%
  dplyr::select(
    cohort,
    predictor,
    label
  )

tab_wide <- tab_ready %>%
  tidyr::pivot_wider(
    names_from = cohort,
    values_from = label
  ) %>%
  dplyr::arrange(
    predictor
  )

kableExtra::kable(
  tab_wide, 
  caption = "Associations between clinical factors and confirmed CKD from the multivariable Bayesian logistic regression model. Odds ratios (ORs) are shown with 95% credible intervals. For screened and unscreened patients, ORs are conditional estimates from the interaction model. For the Enriched cohort, two summary measures are presented: the population-average conditional OR, calculated as the weighted average of screened- and unscreened-specific log-ORs using their observed proportions in the cohort, and the marginal OR, obtained by posterior standardisation of predicted probabilities across all patients in the Enriched cohort. Age effects are expressed for a 25-year difference; the marginal age OR compares 75 versus 50 years"
  )
```



#### Visualisation, screening + conditional enriched

```{r, fig.height=5, fig.width=6.5}
plotac <- "fig_3"
path <- "gitignore/figures"

fig <- data_forest %>%
  dplyr::filter(
    cohort != "unscreened",
    cohort != "enriched_marginal"
  ) %>%
  dplyr::mutate(
    predictor = dplyr::recode(
      predictor,
      "age_25y" = "Age (per 25y)",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_hypertension" = "Arterial hypertension",
      "HF_or_AF" = "Heart failure or atrial fibrillation",
      "FU_atherosclerosis" = "ASCVD",
      "FU_steatosis" = "Liver steatosis",
      "sex_female" = "Female"
    ),
    predictor = factor(
      predictor,
      levels = rev(unique(predictor))
    ),
    cohort = factor(
      cohort,
      levels = c(
        "unscreened",
        "screening",
        "enriched_std_cond",
        "enriched_marginal"
      ),
      labels = c(
        "Unscreened patients",
        "Screening cohort",
        "Enriched cohort (conditional)",
        "Enriched cohort (marginal)"
      )
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = OR,
      y = predictor,
      color = cohort
    )
  ) +
  ggplot2::geom_point(
    position = ggplot2::position_dodge(width = 0.6),
    size = 2
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(
      xmin = CI_L,
      xmax = CI_U
    ),
    position = ggplot2::position_dodge(width = 0.6),
    height = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "gray40"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  ggplot2::scale_color_manual(
    values = c(
      "Unscreened patients" = "#1b9e77",
      "Screening cohort" = "#d95f02",
      "Enriched cohort (conditional)" = "skyblue3",
      "Enriched cohort (marginal)" = "#984ea3"
    )
  ) +
  ggplot2::labs(
    x = "Adjusted odds ratio (log scale) for confirmed CKD",
    y = NULL,
    color = NULL
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(nrow = 1)
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    axis.text.y = ggplot2::element_text(size = 11),
    legend.text = ggplot2::element_text(size = 10)
  )

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggplot2::ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 6.5,
    height = 5
  )
}
```

#### Visualisation, full

```{r, fig.height=5, fig.width=6.5}

plotac <- "suppl_fig_3"
path <- "gitignore/figures"

fig <- data_forest %>%
  dplyr::mutate(
    predictor = dplyr::recode(
      predictor,
      "age_25y" = "Age (per 25y)",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_hypertension" = "Arterial hypertension",
      "HF_or_AF" = "Heart failure or atrial fibrillation",
      "FU_atherosclerosis" = "ASCVD",
      "FU_steatosis" = "Liver steatosis",
      "sex_female" = "Female"
    ),
    predictor = factor(
      predictor,
      levels = rev(unique(predictor))
    ),
    cohort = factor(
      cohort,
      levels = c(
        "unscreened",
        "screening",
        "enriched_std_cond",
        "enriched_marginal"
      ),
      labels = c(
        "Unscreened patients",
        "Screening cohort",
        "Enriched cohort (conditional)",
        "Enriched cohort (marginal)"
      )
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = OR,
      y = predictor,
      color = cohort
    )
  ) +
  ggplot2::geom_point(
    position = ggplot2::position_dodge(width = 0.6),
    size = 2
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(
      xmin = CI_L,
      xmax = CI_U
    ),
    position = ggplot2::position_dodge(width = 0.6),
    height = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "gray40"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  ggplot2::scale_color_manual(
    values = c(
      "Unscreened patients" = "#1b9e77",
      "Screening cohort" = "#d95f02",
      "Enriched cohort (conditional)" = "skyblue3",
      "Enriched cohort (marginal)" = "#984ea3"
    )
  ) +
  ggplot2::labs(
    x = "Adjusted odds ratio (log scale) for confirmed CKD",
    y = NULL,
    color = NULL
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(nrow = 2)
  ) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    legend.position = "top",
    axis.text.y = ggplot2::element_text(size = 11),
    legend.text = ggplot2::element_text(size = 10)
  )

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggplot2::ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 6.5,
    height = 5
  )
}

```

#### Sensitivity analysis model

Fit

```{r}

priors_sens <- c(
  prior(student_t(3, 0, 100), class = "Intercept"),

  prior(normal(0, 5), class = "b", coef = "BMI_over30"),
  prior(normal(0, 5), class = "b", coef = "transplant"),
  prior(normal(0, 5), class = "b", coef = "sex_female"),
  prior(normal(0, 5), class = "b", coef = "FU_DM1"),
  prior(normal(0, 5), class = "b", coef = "FU_DM2"),
  prior(normal(0, 5), class = "b", coef = "FU_atherosclerosis"),
  prior(normal(0, 5), class = "b", coef = "FU_steatosis"),
  prior(normal(0, 5), class = "b", coef = "FU_hypertension"),
  prior(normal(0, 5), class = "b", coef = "HF_or_AF"),
  prior(normal(0, 5), class = "b", coef = "consent1"),

  prior(normal(0, 1), class = "b",
        coef = "transplant:consent1"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:FU_DM1"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:FU_DM2"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:sex_female"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:FU_steatosis"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:FU_hypertension"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:FU_atherosclerosis"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:HF_or_AF"),
  prior(normal(0, 1), class = "b",
        coef = "consent1:BMI_over30")
)

model_sens_lin <- run(
  expr = brm(
    identified_CKD_by_lab ~ 
      s(I( (age-62) / 30), by = consent, k = 5) +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      sex_female * consent + 
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent +
      BMI_over30 * consent,
    data = data_ckd |> 
      dplyr::filter(!is.na(BMI_over30)) |>
      dplyr::mutate(consent = factor(consent)),
    family = bernoulli(link = "logit"),
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = priors_sens,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_sens_lin",
  reuse = TRUE
)

prior_summary(model_sens_lin)
summary(model_sens_lin, robust = TRUE)

exp(fixef(model_sens_lin, robust = TRUE))[,c(1, 3, 4)]
```

Extract posterior and check results

```{r}
dummy_base <- data_ckd_complete[1, ] %>%
  dplyr::mutate(
    transplant = 0,
    FU_DM1 = 0,
    FU_DM2 = 0,
    sex_female = 0,
    FU_steatosis = 0,
    FU_hypertension = 0,
    FU_atherosclerosis = 0,
    HF_or_AF = 0,
    BMI_over30 = 0
  )

d_c0_a50 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(0, levels = c("0", "1")),
    age = 50
  )

d_c0_a75 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(0, levels = c("0", "1")),
    age = 75
  )

lin_c0_a50 <- posterior_linpred(
  model_sens_lin,
  newdata = d_c0_a50
)

lin_c0_a75 <- posterior_linpred(
  model_sens_lin,
  newdata = d_c0_a75
)

age_25y_c0 <- as.numeric(lin_c0_a75 - lin_c0_a50)

d_c1_a50 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(1, levels = c("0", "1")),
    age = 50
  )

d_c1_a75 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(1, levels = c("0", "1")),
    age = 75
  )

lin_c1_a50 <- posterior_linpred(
  model_sens_lin,
  newdata = d_c1_a50
)

lin_c1_a75 <- posterior_linpred(
  model_sens_lin,
  newdata = d_c1_a75
)

age_25y_c1 <- as.numeric(lin_c1_a75 - lin_c1_a50)

post <- as_draws_df(
  model_sens_lin
)

b_all <- post %>%
  dplyr::select(
    dplyr::starts_with("b_")
  )

eff_consent0 <- dplyr::tibble(
  age_25y = age_25y_c0,
  sex_female = b_all$b_sex_female,
  transplant = b_all$b_transplant,
  FU_DM1 = b_all$b_FU_DM1,
  FU_DM2 = b_all$b_FU_DM2,
  FU_steatosis = b_all$b_FU_steatosis,
  FU_hypertension = b_all$b_FU_hypertension,
  FU_atherosclerosis = b_all$b_FU_atherosclerosis,
  HF_or_AF = b_all$b_HF_or_AF,
  BMI_over30 = b_all$b_BMI_over30
)

eff_consent1 <- dplyr::tibble(
  age_25y = age_25y_c1,
  sex_female = b_all$b_sex_female + b_all$`b_consent1:sex_female`,
  transplant = b_all$b_transplant + b_all$`b_transplant:consent1`,
  FU_DM1 = b_all$b_FU_DM1 + b_all$`b_consent1:FU_DM1`,
  FU_DM2 = b_all$b_FU_DM2 + b_all$`b_consent1:FU_DM2`,
  FU_steatosis = b_all$b_FU_steatosis + b_all$`b_consent1:FU_steatosis`,
  FU_hypertension = b_all$b_FU_hypertension + b_all$`b_consent1:FU_hypertension`,
  FU_atherosclerosis = b_all$b_FU_atherosclerosis + b_all$`b_consent1:FU_atherosclerosis`,
  HF_or_AF = b_all$b_HF_or_AF + b_all$`b_consent1:HF_or_AF`,
  BMI_over30 = b_all$b_BMI_over30 + b_all$`b_consent1:BMI_over30`
)

eff_consent_mean <- dplyr::tibble(
  age_25y = age_25y_c0 * (1 - consent_prop) + 
    age_25y_c1 * consent_prop,
  sex_female = eff_consent0$sex_female * (1 - consent_prop) + 
    eff_consent1$sex_female * consent_prop,
  transplant = eff_consent0$transplant * (1 - consent_prop) + 
    eff_consent1$transplant * consent_prop,
  FU_DM1 = eff_consent0$FU_DM1 * (1 - consent_prop) + 
    eff_consent1$FU_DM1 * consent_prop,
  FU_DM2 = eff_consent0$FU_DM2 * (1 - consent_prop) + 
    eff_consent1$FU_DM2 * consent_prop,
  FU_steatosis = eff_consent0$FU_steatosis * (1 - consent_prop) + 
    eff_consent1$FU_steatosis * consent_prop,
  FU_hypertension = eff_consent0$FU_hypertension * (1 - consent_prop) + 
    eff_consent1$FU_hypertension * consent_prop,
  FU_atherosclerosis = eff_consent0$FU_atherosclerosis * (1 - consent_prop) + 
    eff_consent1$FU_atherosclerosis * consent_prop,
  HF_or_AF = eff_consent0$HF_or_AF * (1 - consent_prop) + 
    eff_consent1$HF_or_AF * consent_prop,
  BMI_over30 = eff_consent0$BMI_over30 * (1 - consent_prop) + 
    eff_consent1$BMI_over30 * consent_prop
)

calc_marg_log_or <- function(model, data, var_name, val1, val0) {
  d1 <- data
  
  d1[[var_name]] <- val1
  
  p1 <- posterior_epred(
    model,
    newdata = d1
  )
  
  p1_marg <- rowMeans(p1)

  d0 <- data
  
  d0[[var_name]] <- val0
  
  p0 <- posterior_epred(
    model,
    newdata = d0
  )
  
  p0_marg <- rowMeans(p0)

  odds1 <- p1_marg / (1 - p1_marg)
  
  odds0 <- p0_marg / (1 - p0_marg)

  marg_or <- odds1 / odds0
  
  return(log(marg_or))
}

data_dm_ref <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 0
  )

p0_dm <- posterior_epred(
  model_sens_lin,
  newdata = data_dm_ref
) %>%
  rowMeans()

odds0_dm <- p0_dm / (1 - p0_dm)

data_dm1 <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 1,
    FU_DM2 = 0
  )

p1_dm1 <- posterior_epred(
  model_sens_lin,
  newdata = data_dm1
) %>%
  rowMeans()

odds1_dm1 <- p1_dm1 / (1 - p1_dm1)

data_dm2 <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 1
  )

p1_dm2 <- posterior_epred(
  model_sens_lin,
  newdata = data_dm2
) %>%
  rowMeans()

odds1_dm2 <- p1_dm2 / (1 - p1_dm2)

log_or_dm1_marg <- log(odds1_dm1 / odds0_dm)

log_or_dm2_marg <- log(odds1_dm2 / odds0_dm)

eff_marginal <- dplyr::tibble(
  age_25y = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "age",
    75,
    50
  ),
  sex_female = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "sex_female",
    1,
    0
  ),
  transplant = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "transplant",
    1,
    0
  ),
  FU_DM1 = log_or_dm1_marg,
  FU_DM2 = log_or_dm2_marg,
  FU_steatosis = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "FU_steatosis",
    1,
    0
  ),
  FU_hypertension = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "FU_hypertension",
    1,
    0
  ),
  FU_atherosclerosis = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "FU_atherosclerosis",
    1,
    0
  ),
  HF_or_AF = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "HF_or_AF",
    1,
    0
  ),
  BMI_over30 = calc_marg_log_or(
    model_sens_lin,
    data_ckd_complete,
    "BMI_over30",
    1,
    0
  )
)

summ <- dplyr::bind_rows(
  summarise_draws(
    eff_consent0,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "unscreened"
    ),
  summarise_draws(
    eff_consent1,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "screening"
    ),
  summarise_draws(
    eff_consent_mean,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_std_cond"
    ),
  summarise_draws(
    eff_marginal,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_marginal"
    )
) %>%
  dplyr::rename(
    predictor = variable,
    CI_L = `2.5%`,
    CI_U = `97.5%`
  ) %>%
  dplyr::mutate(
    OR = exp(median),
    CI_L = exp(CI_L),
    CI_U = exp(CI_U)
  )

tab_ready <- summ %>%
  dplyr::mutate(
    label = sprintf(
      "%.2f [%.2f, %.2f]",
      OR,
      CI_L,
      CI_U
    )
  ) %>%
  dplyr::select(
    cohort,
    predictor,
    label
  )

tab_wide <- tab_ready %>%
  tidyr::pivot_wider(
    names_from = cohort,
    values_from = label
  ) %>%
  dplyr::arrange(
    predictor
  )

kableExtra::kable(
  tab_wide,
  caption = "Associations between clinical factors and confirmed CKD from the multivariable Bayesian logistic regression model. Odds ratios (ORs) are shown with 95% credible intervals. For screened and unscreened patients, ORs are conditional estimates from the interaction model. For the Enriched cohort, two summary measures are presented: the population-average conditional OR, calculated as the weighted average of screened- and unscreened-specific log-ORs using their observed proportions in the cohort, and the marginal OR, obtained by posterior standardisation of predicted probabilities across all patients in the Enriched cohort. Age effects are expressed for a 25-year difference comparing 75 versus 50 years."
)
```


## Identified CKD - age-calibrated

### Unadjsuted effects

#### Fit models

```{r}
kableExtra::kable(
  cbind(
    unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          sex_female = if_else(sex == "F", 1, 0),
          BMI10_last = (BMI_last/10),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0)
        ) %>% dplyr::filter(
          consent == 0
        ),
      outcome = "identified_CKD_by_lab_AgeSpecCut",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>% rename(
             `OR [95% CI]` = `OR`
             ),
    unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          sex_female = if_else(sex == "F", 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 1
        ),
      outcome = "identified_CKD_by_lab_AgeSpecCut",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>%
      dplyr::select(OR, N) %>% 
      rename(
             `OR [95% CI]` = `OR`
             ), 
  unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ),
      outcome = "identified_CKD_by_lab_AgeSpecCut",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>%
      dplyr::select(OR, N) %>% 
      rename(
             `OR [95% CI]` = `OR`
             )),
  align = c("r", "c", "c", "c", "c", "c", "c")
) %>%
  add_header_above(
    c(" " = 1, "unscreened" = 2, "screening" = 2, "enriched" = 2)
  ) %>%
  column_spec(1, width = "10em") %>%
  column_spec(c(2, 4, 6), width = "8em") %>%
  column_spec(c(3, 5, 7), width = "5em")

```

#### Visualisation

```{r, fig.height=6.302, fig.width=6.02}

 unadj_eff_con0 <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 0
        ),
      outcome = "identified_CKD_by_lab_AgeSpecCut",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )

 unadj_eff_con1 <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 1
        ),
      outcome = "identified_CKD_by_lab_AgeSpecCut",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )
 
  unadj_eff_overall <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ),
      outcome = "identified_CKD_by_lab_AgeSpecCut",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )
 
df_plot <- bind_rows(
  unadj_eff_con1 %>% mutate(consent = "screening"),
  unadj_eff_overall %>% mutate(consent = "overall")
) %>%
  mutate(
    predictor = recode(
      predictor,
      "age_30y" = "Age (per 30y)",
      "age_60plus" = "Age ≥60y",
      "sex_female" = "Female sex",
      "BMI10_last" = "BMI (per 10 kg/m²)",
      "BMI_over30" = "BMI ≥30 kg/m²",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_DM" = "Any diabetes",
      "FU_atherosclerosis" = "ASCVD",
      "FU_stroke" = "Stroke",
      "FU_PAD" = "Peripheral artery disease",
      "ischemic_heart_disease" = "Ischemic heart disease",
      "FU_hypertension" = "Arterial hypertension",
      "FU_atrial_fibrillation" = "Atrial fibrilation",
      "FU_heart_failure" = "Heart failure",
      "HF_or_AF" = "Heart failure or atrial fibrilation",
      "log2_FU_charlson_index" = "log₂ (Charlson index +1)",
      "charlson_index_5" = "Charlson index ≥5",
      "FU_steatosis" = "Liver steatosis"
    ),
    predictor = factor(predictor, levels = rev(unique(predictor))),
    consent = factor(
  consent,
  levels = c("screening", "overall", "enriched"),
  labels = c("Screening cohort", "Enriched cohort", "No consent cohort")
)
  )
    
  
ggplot(df_plot, aes(x = OR, y = predictor, color = consent, group = consent)) +
  geom_point(
    position = position_dodge(width = 0.6),
    size = 2
    ) +
  geom_errorbarh(
    aes(xmin = CI_L, xmax = CI_U),
    position = position_dodge(width = 0.6),
    height = 0.3
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray40") +
  scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  scale_color_manual(
    values = c(
      "Enriched cohort" = "skyblue3",
      "Screening cohort" = "#d95f02",
      "No consent cohort" = "#1b9e77"
    ), 
    breaks = c(
      "Enriched cohort", "Screening cohort", "No consent cohort")
  ) +
  labs(
    x = "Unadjusted odds ratio (log2 scale) for age-calibrated CKD",
    y = NULL,
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    axis.text.y = element_text(size = 11)
  )

```

### Simple prevalence model

#### Prior

```{r}
prior_intercept <- prior_string(
      paste0("student_t(3, ", out_mean, ", 100)"), 
      class = "Intercept"
    )
```

#### Fit simple mdoel

```{r}
model_simple_AgeCal <- run(
  expr = brm(
    identified_CKD_by_lab_AgeSpecCut ~ +1,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    prior = prior_intercept,
    cores = 1, chains = 2,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_simple_AgeCal",
  reuse = TRUE
)

prior_summary(model_simple_AgeCal)
summary(model_simple_AgeCal, robust = TRUE)

kableExtra::kable(
  inv_logit(fixef(model_simple_AgeCal, robust = TRUE)[, c(1, 3,4)])
)
```

### Consent-specific

```{r}

model_consent_spec_AgeCal <- run(
  expr = brm(
    identified_CKD_by_lab_AgeSpecCut ~ consent,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = prior_intercept,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_consent_spec_AgeCal",
  reuse = TRUE
)

prior_summary(model_consent_spec_AgeCal)
summary(model_consent_spec_AgeCal, robust = TRUE)

posterior <- as_draws_df(model_consent_spec_AgeCal) %>% 
  mutate(
    consent_prev = inv_logit(b_Intercept + b_consent),
    enriched_prev = inv_logit(b_Intercept)
    )

tr <- posterior_epred(model_consent_spec_AgeCal)
tr2 <- rowMeans(tr)

kableExtra::kable(
  data.frame(
    rbind(
      inv_logit(fixef(model_simple_AgeCal, robust = TRUE)[, c(1, 3, 4)]),
      quantile(tr2, probs = c(0.5, 0.025, 0.975)),
      quantile(posterior$enriched_prev, probs = c(0.5, 0.025, 0.975)),
      quantile(posterior$consent_prev, probs = c(0.5, 0.025, 0.975))
    )
  ) %>%
    dplyr::mutate(
      Estimand = c(
        "Marginal prevalence - simple model",
        "Marginal prevalence - consent-adj. model",
        "Prevalence - consent=0",
        "Prevalence - consent=1"
      )
    ) %>% dplyr::select(
      Estimand, Estimate, Q2.5, Q97.5
    )
)
```

### Non-linear time model

#### Fit model

```{r}
model_time_nl_AgeCal <- run(
  expr = brm(
    identified_CKD_by_lab_AgeSpecCut ~
      consent +
      s(I( (age-62) / 30), by = consent, k = 5),
    family = bernoulli(link = "logit"),
    data = data_ckd %>% dplyr::mutate(consent = factor(consent)),
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = prior_intercept,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_time_nl_AgeCal",
  reuse = TRUE
)

prior_summary(model_time_nl_AgeCal)
summary(model_time_nl_AgeCal, robust = TRUE)

```

#### Get posteriors

```{r}
quantile(data_ckd$age, probs = c(0, 0.01, 0.05, 0.25, 0.5, 0.75, 0.95, 0.99))

age_seq <- c(
  seq(min(data_ckd$age),
    max(data_ckd$age),
    length.out = 101
  ),
  c(seq(25, 80, by = 5))
)

consent_prop <- mean(data_ckd$consent)

new0 <- data.frame(
  age = age_seq,
  consent = factor(0, levels = c(0, 1))
)

new1 <- data.frame(
  age = age_seq,
  consent = factor(1, levels = c(0, 1))
)

p0 <- posterior_epred(model_time_nl_AgeCal, newdata = new0)
p1 <- posterior_epred(model_time_nl_AgeCal, newdata = new1)
p_overall <- (1 - consent_prop) * p0 + consent_prop * p1

summarize_epred <- function(mat, cohort) {
  posterior_summary(mat, robust = TRUE) %>%
    as.data.frame() %>%
    dplyr::mutate(
      cohort = cohort,
      age = age_seq
    )
}

prediction_NL_AgeCal <- dplyr::bind_rows(
  summarize_epred(p1, "screening"),
  summarize_epred(p0, "unscreened"),
  summarize_epred(p_overall, "enriched")
  ) %>%
  dplyr::mutate(age_calibrated = "Age calibrated")

kableExtra::kable(
  prediction_NL_AgeCal %>%
    dplyr::filter(age %in% seq(25, 80, by = 5)) %>% 
    dplyr::select(-age_calibrated)
)
```

#### Visualisation

```{r}
prediction_NL_AgeCal %>% 
  dplyr::filter(cohort != 'enriched') %>% 
  ggplot(
    aes(
      x = age,
      y = Estimate,
      color = cohort,
      fill = cohort
    )
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = Q2.5,
      ymax = Q97.5
    ),
    alpha = 0.4,
    color = NA
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)
  ) +
  labs(
    x = "Age (years)",
    y = "Estimated probability of CKD new case"
  ) +
  scale_color_manual(
    values = c(
        "unscreened" = "#1b9e77",
        "screening" = "#d95f02",
        "enriched" = "skyblue3"
    ),
    name = "Cohort"
  ) +
  scale_fill_manual(
     values = c(
        "unscreened" = "#1b9e77",
        "screening" = "#d95f02",
        "enriched" = "skyblue3"
    ),
    name = "Cohort"
  ) +
  facet_grid(rows = vars(cohort)) +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    strip.text.x = element_text(size = 12),
    legend.position = "none"
  )
```

#### Joint visualisation -- enriched cohort

```{r, fig.width=5.5, fig.height=7}

prediction <- bind_rows(prediction_NL, prediction_NL_AgeCal)

prediction %>% 
  dplyr::filter(cohort == 'enriched') %>% 
  ggplot(
    aes(
      x = age,
      y = Estimate,
      color = age_calibrated,
      fill = age_calibrated
    )
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = Q2.5,
      ymax = Q97.5
    ),
    alpha = 0.4,
    color = NA
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)
  ) +
  labs(
    x = "Age (years)",
    y = "Estimated probability of confirmed CKD"
  ) +
  scale_color_manual(
    values = c(
        "No calibration" = "pink3",
        "Age calibrated" = "cyan3"
    ),
    name = "Age calibration"
  ) +
  scale_fill_manual(
     values = c(
        "No calibration" = "pink3",
        "Age calibrated" = "cyan3"
    ),
    name = "Age calibration"
  ) +
  facet_grid(rows = vars(age_calibrated)) +
  theme(
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    legend.position = "none"
  )
```

#### Joint visualisation -- screening cohort

```{r, fig.width=5.5, fig.height=7}
plotac <- "suppl_fig_6"
path <- "gitignore/figures"

prediction <- bind_rows(
  prediction_NL,
  prediction_NL_AgeCal
)

fig <- prediction %>%
  dplyr::filter(cohort == "screening") %>%
  ggplot(
    aes(
      x = age,
      y = Estimate,
      color = age_calibrated,
      fill = age_calibrated
    )
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = Q2.5,
      ymax = Q97.5
    ),
    alpha = 0.4,
    color = NA
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)
  ) +
  labs(
    x = "Age (years)",
    y = "Estimated probability of confirmed CKD"
  ) +
  scale_color_manual(
    values = c(
      "No calibration" = "pink3",
      "Age calibrated" = "cyan3"
    ),
    name = "Age calibration"
  ) +
  scale_fill_manual(
    values = c(
      "No calibration" = "pink3",
      "Age calibrated" = "cyan3"
    ),
    name = "Age calibration"
  ) +
  facet_grid(
    rows = vars(age_calibrated)
  ) +
  theme(
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    legend.position = "none"
  )

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 5.5,
    height = 7
  )
}
```

### Risk factors model

#### Can age be lienarized?

```{r}
m_tr <- mgcv::gam(
  identified_CKD_by_lab_AgeSpecCut ~ 
      s(age,  k = 5) +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent,
  data = data_ckd,
  family = binomial(link = 'logit')
  )

summary(m_tr)
plot(m_tr)
```

#### Fit model

```{r}


model_complex_lin_AgeCal <- run(
  expr = brm(
    identified_CKD_by_lab_AgeSpecCut ~ 
      I( (age-62) / 30) * consent +
      transplant * consent +
      sex_female * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = priors,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_complex_lin_AgeCal",
  reuse = TRUE
)

prior_summary(model_complex_lin_AgeCal)
summary(model_complex_lin_AgeCal, robust = TRUE)

exp(fixef(model_complex_lin_AgeCal, robust = TRUE))[,c(1, 3, 4)]
```

#### Extract posterior

```{r}
consent_prop <- mean(data_ckd$consent)

age_scale <- 25 / 30

post <- posterior::as_draws_df(
  model_complex_lin_AgeCal
)

b_all <- post %>%
  dplyr::select(
    dplyr::starts_with("b_")
  )

base <- b_all %>%
  dplyr::transmute(
    Intercept = b_Intercept,
    age_25y = b_IageM62D30 * age_scale,
    sex_female = b_sex_female,
    transplant = b_transplant,
    FU_DM1 = b_FU_DM1,
    FU_DM2 = b_FU_DM2,
    FU_steatosis = b_FU_steatosis,
    FU_hypertension = b_FU_hypertension,
    FU_atherosclerosis = b_FU_atherosclerosis,
    HF_or_AF = b_HF_or_AF
  )

eff_consent0 <- base

eff_consent1 <- base %>%
  dplyr::mutate(
    Intercept = Intercept + b_all$b_consent,
    age_25y = age_25y + b_all$`b_IageM62D30:consent` * age_scale,
    transplant = transplant + b_all$`b_consent:transplant`,
    FU_DM1 = FU_DM1 + b_all$`b_consent:FU_DM1`,
    FU_DM2 = FU_DM2 + b_all$`b_consent:FU_DM2`,
    FU_steatosis = FU_steatosis + b_all$`b_consent:FU_steatosis`,
    FU_hypertension = FU_hypertension + b_all$`b_consent:FU_hypertension`,
    FU_atherosclerosis = FU_atherosclerosis + b_all$`b_consent:FU_atherosclerosis`,
    HF_or_AF = HF_or_AF + b_all$`b_consent:HF_or_AF`,
    sex_female = sex_female + b_all$`b_consent:sex_female`
  )

eff_consent_mean <- base %>%
  dplyr::mutate(
    Intercept = Intercept + b_all$b_consent * consent_prop,
    age_25y = age_25y + b_all$`b_IageM62D30:consent` * consent_prop * age_scale,
    transplant = transplant + b_all$`b_consent:transplant` * consent_prop,
    FU_DM1 = FU_DM1 + b_all$`b_consent:FU_DM1` * consent_prop,
    FU_DM2 = FU_DM2 + b_all$`b_consent:FU_DM2` * consent_prop,
    FU_steatosis = FU_steatosis + b_all$`b_consent:FU_steatosis` * consent_prop,
    FU_hypertension = FU_hypertension + b_all$`b_consent:FU_hypertension` * consent_prop,
    FU_atherosclerosis = FU_atherosclerosis + b_all$`b_consent:FU_atherosclerosis` * consent_prop,
    HF_or_AF = HF_or_AF + b_all$`b_consent:HF_or_AF` * consent_prop,
    sex_female = sex_female + b_all$`b_consent:sex_female` * consent_prop
  )

data_dm_ref <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 0
  )

p0_dm <- brms::posterior_epred(
  model_complex_lin_AgeCal,
  newdata = data_dm_ref
) %>%
  rowMeans()

odds0_dm <- p0_dm / (1 - p0_dm)

data_dm1 <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 1,
    FU_DM2 = 0
  )

p1_dm1 <- brms::posterior_epred(
  model_complex_lin_AgeCal,
  newdata = data_dm1
) %>%
  rowMeans()

odds1_dm1 <- p1_dm1 / (1 - p1_dm1)

data_dm2 <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 1
  )

p1_dm2 <- brms::posterior_epred(
  model_complex_lin_AgeCal,
  newdata = data_dm2
) %>%
  rowMeans()

odds1_dm2 <- p1_dm2 / (1 - p1_dm2)

log_or_dm1_marg <- log(odds1_dm1 / odds0_dm)
log_or_dm2_marg <- log(odds1_dm2 / odds0_dm)

eff_marginal <- dplyr::tibble(
  age_25y = calc_marg_log_or(
    model_complex_lin_AgeCal,
    data_ckd,
    "age",
    75,
    50
  ),
  sex_female = calc_marg_log_or(
    model_complex_lin_AgeCal,
    data_ckd,
    "sex_female",
    1,
    0
  ),
  transplant = calc_marg_log_or(
    model_complex_lin_AgeCal,
    data_ckd,
    "transplant",
    1,
    0
  ),
  FU_DM1 = log_or_dm1_marg,
  FU_DM2 = log_or_dm2_marg,
  FU_steatosis = calc_marg_log_or(
    model_complex_lin_AgeCal,
    data_ckd,
    "FU_steatosis",
    1,
    0
  ),
  FU_hypertension = calc_marg_log_or(
    model_complex_lin_AgeCal,
    data_ckd,
    "FU_hypertension",
    1,
    0
  ),
  FU_atherosclerosis = calc_marg_log_or(
    model_complex_lin_AgeCal,
    data_ckd,
    "FU_atherosclerosis",
    1,
    0
  ),
  HF_or_AF = calc_marg_log_or(
    model_complex_lin_AgeCal,
    data_ckd,
    "HF_or_AF",
    1,
    0
  )
)

summ <- dplyr::bind_rows(
  posterior::summarise_draws(
    eff_consent0,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "unscreened"
    ),
  posterior::summarise_draws(
    eff_consent1,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "screening"
    ),
  posterior::summarise_draws(
    eff_consent_mean,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "enriched_std_cond"
    ),
  posterior::summarise_draws(
    eff_marginal,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "enriched_marginal"
    )
) %>%
  dplyr::rename(
    predictor = variable,
    CI_L = `2.5%`,
    CI_U = `97.5%`
  ) %>%
  dplyr::mutate(
    OR = exp(median),
    CI_L = exp(CI_L),
    CI_U = exp(CI_U)
  )

data_forest <- summ %>%
  dplyr::filter(
    predictor != "Intercept"
  ) %>%
  dplyr::mutate(
    predictor = factor(
      predictor,
      levels = rev(
        unique(
          predictor
        )
      )
    )
  )

tab_ready <- summ %>%
  dplyr::filter(
    predictor != "Intercept"
  ) %>%
  dplyr::mutate(
    label = sprintf(
      "%.2f [%.2f, %.2f]",
      OR,
      CI_L,
      CI_U
    )
  ) %>%
  dplyr::select(
    cohort,
    predictor,
    label
  )

tab_wide <- tab_ready %>%
  tidyr::pivot_wider(
    names_from = cohort,
    values_from = label
  ) %>%
  dplyr::arrange(
    predictor
  )

kableExtra::kable(tab_wide, caption = "Associations between clinical factors and age-calibrated CKD from the multivariable Bayesian logistic regression model. Odds ratios (ORs) are shown with 95% credible intervals. For screened and unscreened patients, ORs are conditional estimates from the interaction model. For the Enriched cohort, two summary measures are presented: the population-average conditional OR, calculated as the weighted average of screened- and unscreened-specific log-ORs using their observed proportions in the cohort, and the marginal OR, obtained by posterior standardisation of predicted probabilities across all patients in the Enriched cohort. Age effects are expressed for a 25-year difference; the marginal age OR compares 75 versus 50 years.")

```

#### Visualisation, screening + enriched (conditional)

```{r, fig.height=5, fig.width=6.5}
data_forest %>%
  dplyr::filter(
    cohort != "unscreened",
    cohort != "enriched_marginal"
  ) %>%
  dplyr::mutate(
    predictor = dplyr::recode(
      predictor,
      "age_25y" = "Age (per 25y)",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_hypertension" = "Arterial hypertension",
      "HF_or_AF" = "Heart failure or atrial fibrillation",
      "FU_atherosclerosis" = "ASCVD",
      "FU_steatosis" = "Liver steatosis",
      "sex_female" = "Female"
    ),
    predictor = factor(
      predictor,
      levels = rev(
        unique(
          predictor
        )
      )
    ),
    cohort = factor(
      cohort,
      levels = c(
        "unscreened",
        "screening",
        "enriched_std_cond",
        "enriched_marginal"
      ),
      labels = c(
        "Unscreened patients",
        "Screening cohort",
        "Enriched cohort (conditional)",
        "Enriched cohort (marginal)"
      )
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = OR,
      y = predictor,
      color = cohort
    )
  ) +
  ggplot2::geom_point(
    position = ggplot2::position_dodge(width = 0.6),
    size = 2
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(
      xmin = CI_L,
      xmax = CI_U
    ),
    position = ggplot2::position_dodge(width = 0.6),
    height = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "gray40"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  ggplot2::scale_color_manual(
    values = c(
      "Unscreened patients" = "#1b9e77",
      "Screening cohort" = "#d95f02",
      "Enriched cohort (conditional)" = "skyblue3",
      "Enriched cohort (marginal)" = "#984ea3"
    )
  ) +
  ggplot2::labs(
    x = "Adjusted odds ratio (log scale) for age-calibrated CKD",
    y = NULL,
    color = NULL
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      nrow = 1
    )
  ) +
  ggplot2::theme_minimal(
    base_size = 12
  ) +
  ggplot2::theme(
    legend.position = "top",
    axis.text.y = ggplot2::element_text(size = 11),
    legend.text = ggplot2::element_text(size = 10)
  )
```



#### Visualisation, all

```{r, fig.height=5, fig.width=6.5}
plotac <- "suppl_fig_7"
path <- "gitignore/figures"

fig <- data_forest %>%
  dplyr::mutate(
    predictor = dplyr::recode(
      predictor,
      "age_25y" = "Age (per 25y)",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_hypertension" = "Arterial hypertension",
      "HF_or_AF" = "Heart failure or atrial fibrillation",
      "FU_atherosclerosis" = "ASCVD",
      "FU_steatosis" = "Liver steatosis",
      "sex_female" = "Female"
    ),
    predictor = factor(
      predictor,
      levels = rev(
        unique(
          predictor
        )
      )
    ),
    cohort = factor(
      cohort,
      levels = c(
        "unscreened",
        "screening",
        "enriched_std_cond",
        "enriched_marginal"
      ),
      labels = c(
        "Unscreened patients",
        "Screening cohort",
        "Enriched cohort (conditional)",
        "Enriched cohort (marginal)"
      )
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = OR,
      y = predictor,
      color = cohort
    )
  ) +
  ggplot2::geom_point(
    position = ggplot2::position_dodge(width = 0.6),
    size = 2
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(
      xmin = CI_L,
      xmax = CI_U
    ),
    position = ggplot2::position_dodge(width = 0.6),
    height = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "gray40"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  ggplot2::scale_color_manual(
    values = c(
      "Unscreened patients" = "#1b9e77",
      "Screening cohort" = "#d95f02",
      "Enriched cohort (conditional)" = "skyblue3",
      "Enriched cohort (marginal)" = "#984ea3"
    )
  ) +
  ggplot2::labs(
    x = "Adjusted odds ratio (log scale) for age-calibrated CKD",
    y = NULL,
    color = NULL
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      nrow = 2
    )
  ) +
  ggplot2::theme_minimal(
    base_size = 12
  ) +
  ggplot2::theme(
    legend.position = "top",
    axis.text.y = ggplot2::element_text(size = 11),
    legend.text = ggplot2::element_text(size = 10)
  )

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggplot2::ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 6.5,
    height = 5
  )
}
```

#### Sensitivity analysis

Fit model

```{r}
model_sens_lin_AgeCal <- run(
  expr = brm(
    identified_CKD_by_lab_AgeSpecCut ~ 
      s(I( (age-62) / 30), by = consent, k = 5) +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      sex_female * consent + 
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent +
      BMI_over30 * consent,
    data = data_ckd |> 
      dplyr::filter(!is.na(BMI_over30)) |>
      dplyr::mutate(consent = factor(consent)),
    family = bernoulli(link = "logit"),
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = priors_sens,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_sens_lin_AgeCal",
  reuse = TRUE
)

prior_summary(model_sens_lin_AgeCal)
summary(model_sens_lin_AgeCal, robust = TRUE)

exp(fixef(model_sens_lin_AgeCal, robust = TRUE))[,c(1, 3, 4)]
```

Extract posteriors and show results

```{r}
dummy_base <- data_ckd_complete[1, ] %>%
  dplyr::mutate(
    transplant = 0,
    FU_DM1 = 0,
    FU_DM2 = 0,
    sex_female = 0,
    FU_steatosis = 0,
    FU_hypertension = 0,
    FU_atherosclerosis = 0,
    HF_or_AF = 0,
    BMI_over30 = 0
  )

d_c0_a50 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(0, levels = c("0", "1")),
    age = 50
  )

d_c0_a75 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(0, levels = c("0", "1")),
    age = 75
  )

lin_c0_a50 <- posterior_linpred(
  model_sens_lin_AgeCal,
  newdata = d_c0_a50
)

lin_c0_a75 <- posterior_linpred(
  model_sens_lin_AgeCal,
  newdata = d_c0_a75
)

age_25y_c0 <- as.numeric(lin_c0_a75 - lin_c0_a50)

d_c1_a50 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(1, levels = c("0", "1")),
    age = 50
  )

d_c1_a75 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(1, levels = c("0", "1")),
    age = 75
  )

lin_c1_a50 <- posterior_linpred(
  model_sens_lin_AgeCal,
  newdata = d_c1_a50
)

lin_c1_a75 <- posterior_linpred(
  model_sens_lin_AgeCal,
  newdata = d_c1_a75
)

age_25y_c1 <- as.numeric(lin_c1_a75 - lin_c1_a50)

post <- as_draws_df(
  model_sens_lin_AgeCal
)

b_all <- post %>%
  dplyr::select(
    dplyr::starts_with("b_")
  )

eff_consent0 <- dplyr::tibble(
  age_25y = age_25y_c0,
  sex_female = b_all$b_sex_female,
  transplant = b_all$b_transplant,
  FU_DM1 = b_all$b_FU_DM1,
  FU_DM2 = b_all$b_FU_DM2,
  FU_steatosis = b_all$b_FU_steatosis,
  FU_hypertension = b_all$b_FU_hypertension,
  FU_atherosclerosis = b_all$b_FU_atherosclerosis,
  HF_or_AF = b_all$b_HF_or_AF,
  BMI_over30 = b_all$b_BMI_over30
)

eff_consent1 <- dplyr::tibble(
  age_25y = age_25y_c1,
  sex_female = b_all$b_sex_female + b_all$`b_consent1:sex_female`,
  transplant = b_all$b_transplant + b_all$`b_transplant:consent1`,
  FU_DM1 = b_all$b_FU_DM1 + b_all$`b_consent1:FU_DM1`,
  FU_DM2 = b_all$b_FU_DM2 + b_all$`b_consent1:FU_DM2`,
  FU_steatosis = b_all$b_FU_steatosis + b_all$`b_consent1:FU_steatosis`,
  FU_hypertension = b_all$b_FU_hypertension + b_all$`b_consent1:FU_hypertension`,
  FU_atherosclerosis = b_all$b_FU_atherosclerosis + b_all$`b_consent1:FU_atherosclerosis`,
  HF_or_AF = b_all$b_HF_or_AF + b_all$`b_consent1:HF_or_AF`,
  BMI_over30 = b_all$b_BMI_over30 + b_all$`b_consent1:BMI_over30`
)

eff_consent_mean <- dplyr::tibble(
  age_25y = age_25y_c0 * (1 - consent_prop) + 
    age_25y_c1 * consent_prop,
  sex_female = eff_consent0$sex_female * (1 - consent_prop) + 
    eff_consent1$sex_female * consent_prop,
  transplant = eff_consent0$transplant * (1 - consent_prop) + 
    eff_consent1$transplant * consent_prop,
  FU_DM1 = eff_consent0$FU_DM1 * (1 - consent_prop) + 
    eff_consent1$FU_DM1 * consent_prop,
  FU_DM2 = eff_consent0$FU_DM2 * (1 - consent_prop) + 
    eff_consent1$FU_DM2 * consent_prop,
  FU_steatosis = eff_consent0$FU_steatosis * (1 - consent_prop) + 
    eff_consent1$FU_steatosis * consent_prop,
  FU_hypertension = eff_consent0$FU_hypertension * (1 - consent_prop) + 
    eff_consent1$FU_hypertension * consent_prop,
  FU_atherosclerosis = eff_consent0$FU_atherosclerosis * (1 - consent_prop) + 
    eff_consent1$FU_atherosclerosis * consent_prop,
  HF_or_AF = eff_consent0$HF_or_AF * (1 - consent_prop) + 
    eff_consent1$HF_or_AF * consent_prop,
  BMI_over30 = eff_consent0$BMI_over30 * (1 - consent_prop) + 
    eff_consent1$BMI_over30 * consent_prop
)

calc_marg_log_or <- function(model, data, var_name, val1, val0) {
  d1 <- data
  
  d1[[var_name]] <- val1
  
  p1 <- posterior_epred(
    model,
    newdata = d1
  )
  
  p1_marg <- rowMeans(p1)

  d0 <- data
  
  d0[[var_name]] <- val0
  
  p0 <- posterior_epred(
    model,
    newdata = d0
  )
  
  p0_marg <- rowMeans(p0)

  odds1 <- p1_marg / (1 - p1_marg)
  
  odds0 <- p0_marg / (1 - p0_marg)

  marg_or <- odds1 / odds0
  
  return(log(marg_or))
}

data_dm_ref <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 0
  )

p0_dm <- posterior_epred(
  model_sens_lin_AgeCal,
  newdata = data_dm_ref
) %>%
  rowMeans()

odds0_dm <- p0_dm / (1 - p0_dm)

data_dm1 <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 1,
    FU_DM2 = 0
  )

p1_dm1 <- posterior_epred(
  model_sens_lin_AgeCal,
  newdata = data_dm1
) %>%
  rowMeans()

odds1_dm1 <- p1_dm1 / (1 - p1_dm1)

data_dm2 <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 1
  )

p1_dm2 <- posterior_epred(
  model_sens_lin_AgeCal,
  newdata = data_dm2
) %>%
  rowMeans()

odds1_dm2 <- p1_dm2 / (1 - p1_dm2)

log_or_dm1_marg <- log(odds1_dm1 / odds0_dm)

log_or_dm2_marg <- log(odds1_dm2 / odds0_dm)

eff_marginal <- dplyr::tibble(
  age_25y = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "age",
    75,
    50
  ),
  sex_female = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "sex_female",
    1,
    0
  ),
  transplant = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "transplant",
    1,
    0
  ),
  FU_DM1 = log_or_dm1_marg,
  FU_DM2 = log_or_dm2_marg,
  FU_steatosis = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "FU_steatosis",
    1,
    0
  ),
  FU_hypertension = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "FU_hypertension",
    1,
    0
  ),
  FU_atherosclerosis = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "FU_atherosclerosis",
    1,
    0
  ),
  HF_or_AF = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "HF_or_AF",
    1,
    0
  ),
  BMI_over30 = calc_marg_log_or(
    model_sens_lin_AgeCal,
    data_ckd_complete,
    "BMI_over30",
    1,
    0
  )
)

summ <- dplyr::bind_rows(
  summarise_draws(
    eff_consent0,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "unscreened"
    ),
  summarise_draws(
    eff_consent1,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "screening"
    ),
  summarise_draws(
    eff_consent_mean,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_std_cond"
    ),
  summarise_draws(
    eff_marginal,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_marginal"
    )
) %>%
  dplyr::rename(
    predictor = variable,
    CI_L = `2.5%`,
    CI_U = `97.5%`
  ) %>%
  dplyr::mutate(
    OR = exp(median),
    CI_L = exp(CI_L),
    CI_U = exp(CI_U)
  )

tab_ready <- summ %>%
  dplyr::mutate(
    label = sprintf(
      "%.2f [%.2f, %.2f]",
      OR,
      CI_L,
      CI_U
    )
  ) %>%
  dplyr::select(
    cohort,
    predictor,
    label
  )

tab_wide <- tab_ready %>%
  tidyr::pivot_wider(
    names_from = cohort,
    values_from = label
  ) %>%
  dplyr::arrange(
    predictor
  )

kableExtra::kable(
  tab_wide,
  caption = "Associations between clinical factors and confirmed CKD (defined by an age-specific eGFR threshold) from the multivariable Bayesian logistic regression model. Odds ratios (ORs) are shown with 95% credible intervals. For screened and unscreened patients, ORs are conditional estimates from the interaction model. For the Enriched cohort, two summary measures are presented: the population-average conditional OR, calculated as the weighted average of screened- and unscreened-specific log-ORs using their observed proportions in the cohort, and the marginal OR, obtained by posterior standardisation of predicted probabilities across all patients in the Enriched cohort. Age effects are expressed for a 25-year difference comparing 75 versus 50 years."
)
```


## New CKD cases

### Unadjsuted effects

#### Fit models

```{r}
kableExtra::kable(
  cbind(
    unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          sex_female = if_else(sex == "F", 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 0
        ),
      outcome = "new_cases",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>% rename(
             `OR [95% CI]` = `OR`
             ),
    unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),          
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          sex_female = if_else(sex == "F", 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 1
        ),
      outcome = "new_cases",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>%
      dplyr::select(OR, N) %>% 
      rename(
             `OR [95% CI]` = `OR`
             ), 
  unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          BMI10_last = (BMI_last/10)
        ),
      outcome = "new_cases",
      predictors = preds,
      family = "logistic",
      rounding = 2
    ) %>%
      dplyr::select(OR, N) %>% 
      rename(
             `OR [95% CI]` = `OR`
             )),
  align = c("r", "c", "c", "c", "c", "c", "c")
) %>%
  add_header_above(
    c(" " = 1, "unscreened" = 2, "screening" = 2, "enriched" = 2)
  ) %>%
  column_spec(1, width = "10em") %>%
  column_spec(c(2, 4, 6), width = "8em") %>%
  column_spec(c(3, 5, 7), width = "5em")

```

#### Visualisation

```{r, fig.height=6.302, fig.width=6.02}

 unadj_eff_con0 <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 0
        ),
      outcome = "new_cases",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )

 unadj_eff_con1 <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ) %>% dplyr::filter(
          consent == 1
        ),
      outcome = "new_cases",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )
 
  unadj_eff_overall <- unadj_effect(
      dat = data_ckd %>%
        dplyr::mutate(
          age_30y = (age/30),
          log2_FU_charlson_index = log2(FU_charlson_index + 1),
          age_60plus = if_else(age >= 60, 1, 0),
          charlson_index_5 = if_else(FU_charlson_index >= 5, 1, 0),
          BMI10_last = (BMI_last/10)
        ),
      outcome = "new_cases",
      predictors = preds,
      family = "logistic",
      paste_CI = FALSE
    )
 
df_plot <- bind_rows(
  unadj_eff_con1 %>% mutate(consent = "screening"),
  unadj_eff_overall %>% mutate(consent = "enriched")
) %>%
  mutate(
    predictor = recode(
      predictor,
      "age_30y" = "Age (per 30y)",
      "age_60plus" = "Age ≥60y",
      "sex_female" = "Female sex",
      "BMI10_last" = "BMI (per 10 kg/m²)",
      "BMI_over30" = "BMI ≥30 kg/m²",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_DM" = "Any diabetes",
      "FU_atherosclerosis" = "ASCVD",
      "FU_stroke" = "Stroke",
      "FU_PAD" = "Peripheral artery disease",
      "ischemic_heart_disease" = "Ischemic heart disease",
      "FU_hypertension" = "Arterial hypertension",
      "FU_atrial_fibrillation" = "Atrial fibrilation",
      "FU_heart_failure" = "Heart failure",
      "HF_or_AF" = "Heart failure or atrial fibrilation",
      "log2_FU_charlson_index" = "log₂ (Charlson index +1)",
      "charlson_index_5" = "Charlson index ≥5",
      "FU_steatosis" = "Liver steatosis"
    ),
    predictor = factor(predictor, levels = rev(unique(predictor))),
    consent = factor(
  consent,
  levels = c("screening", "enriched", "unscreened"),
  labels = c("Screening cohort", "Enriched cohort", "No consent cohort")
)
  )
    
  
ggplot(df_plot, aes(x = OR, y = predictor, color = consent, group = consent)) +
  geom_point(
    position = position_dodge(width = 0.6),
    size = 2
    ) +
  geom_errorbarh(
    aes(xmin = CI_L, xmax = CI_U),
    position = position_dodge(width = 0.6),
    height = 0.3
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray40") +
  scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  scale_color_manual(
    values = c(
      "Enriched cohort" = "skyblue3",
      "Screening cohort" = "#d95f02",
      "No consent cohort" = "#1b9e77"
    ), 
    breaks = c(
      "Enriched cohort", "Screening cohort", "No consent cohort")
  ) +
  labs(
    x = "Unadjusted odds ratio (log2 scale) for new CKD cases",
    y = NULL,
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    axis.text.y = element_text(size = 11)
  )

```

### Simple prevalence model

#### Prior

```{r}
prior_intercept <- prior_string(
      paste0("student_t(3, ", out_mean, ", 100)"), 
      class = "Intercept"
    )

```

#### Fit simple mdoel

```{r}
model_simple_NewCases <- run(
  expr = brm(
    new_cases ~ +1,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    prior = prior_intercept,
    cores = 1, chains = 2,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_simple_NewCases",
  reuse = TRUE
)

prior_summary(model_simple_NewCases)
summary(model_simple_NewCases, robust = TRUE)

kableExtra::kable(
  inv_logit(fixef(model_simple_NewCases, robust = TRUE)[, c(1, 3,4)])
)
```

### Consent-specific

```{r}

model_consent_spec_NewCases <- run(
  expr = brm(
    new_cases ~ consent,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = prior_intercept,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_consent_spec_NewCases",
  reuse = TRUE
)

prior_summary(model_consent_spec_NewCases)
summary(model_consent_spec_NewCases, robust = TRUE)

posterior <- as_draws_df(model_consent_spec_NewCases) %>% 
  mutate(
    consent_prev = inv_logit(b_Intercept + b_consent),
    enriched_prev = inv_logit(b_Intercept)
    )

tr <- posterior_epred(model_consent_spec_NewCases)
tr2 <- rowMeans(tr)

kableExtra::kable(
  data.frame(
    rbind(
      inv_logit(fixef(model_simple_NewCases, robust = TRUE)[, c(1, 3, 4)]),
      quantile(tr2, probs = c(0.5, 0.025, 0.975)),
      quantile(posterior$enriched_prev, probs = c(0.5, 0.025, 0.975)),
      quantile(posterior$consent_prev, probs = c(0.5, 0.025, 0.975))
    )
  ) %>%
    dplyr::mutate(
      Estimand = c(
        "Marginal prevalence - simple model",
        "Marginal prevalence - consent-adj. model",
        "Prevalence - consent=0",
        "Prevalence - consent=1"
      )
    ) %>% dplyr::select(
      Estimand, Estimate, Q2.5, Q97.5
    )
)
```

### Non-linear time model

#### Fit model

```{r}
model_time_nl_NewCases <- run(
  expr = brm(
    new_cases ~
      consent +
      s(I( (age-62) / 30), by = consent, k = 5),
    family = bernoulli(link = "logit"),
    data = data_ckd %>% dplyr::mutate(consent = factor(consent)),
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = prior_intercept,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_time_nl_NewCases",
  reuse = TRUE
)

prior_summary(model_time_nl_NewCases)
summary(model_time_nl_NewCases, robust = TRUE)

```

#### Get posteriors

```{r}
quantile(data_ckd$age, probs = c(0, 0.01, 0.05, 0.25, 0.5, 0.75, 0.95, 0.99))

age_seq <- c(
  seq(min(data_ckd$age),
    max(data_ckd$age),
    length.out = 101
  ),
  c(seq(25, 80, by = 5))
)

consent_prop <- mean(data_ckd$consent)

new0 <- data.frame(
  age = age_seq,
  consent = factor(0, levels = c(0, 1))
)

new1 <- data.frame(
  age = age_seq,
  consent = factor(1, levels = c(0, 1))
)

p0 <- posterior_epred(model_time_nl_NewCases, newdata = new0)
p1 <- posterior_epred(model_time_nl_NewCases, newdata = new1)
p_overall <- (1 - consent_prop) * p0 + consent_prop * p1

summarize_epred <- function(mat, cohort) {
  posterior_summary(mat, robust = TRUE) %>%
    as.data.frame() %>%
    dplyr::mutate(
      cohort = cohort,
      age = age_seq
    )
}

prediction <- dplyr::bind_rows(
  summarize_epred(p1, "screening"),
  summarize_epred(p0, "unscreened"),
  summarize_epred(p_overall, "enriched")
)

kableExtra::kable(
  prediction %>%
    dplyr::filter(age %in% seq(25, 80, by = 5))
)

```

#### Visualisation

```{r}
prediction %>% 
  dplyr::filter(cohort != 'unscreened') %>% 
  ggplot(
    aes(
      x = age,
      y = Estimate,
      color = cohort,
      fill = cohort
    )
  ) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(
      ymin = Q2.5,
      ymax = Q97.5
    ),
    alpha = 0.4,
    color = NA
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)
  ) +
  labs(
    x = "Age (years)",
    y = "Estimated probability of new CKD case"
  ) +
  scale_color_manual(
    values = c(
        "unscreened" = "#1b9e77",
        "screening" = "#d95f02",
        "enriched" = "skyblue3"
    ),
    name = "Cohort"
  ) +
  scale_fill_manual(
     values = c(
        "unscreened" = "#1b9e77",
        "screening" = "#d95f02",
        "enriched" = "skyblue3"
    ),
    name = "Cohort"
  ) +
  facet_grid(rows = vars(cohort)) +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    strip.text.x = element_text(size = 12),
    legend.position = "none"
  )
```

### Risk factors model

#### Can age be lienarized?

```{r}
m_tr <- mgcv::gam(
  new_cases ~ 
      s(age,  k = 5) +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      sex_female* consent +
      HF_or_AF * consent,
  data = data_ckd,
  family = binomial(link = 'logit')
  )

summary(m_tr)
plot(m_tr)
```

#### Fit

```{r}

model_complex_lin_NewCases <- run(
  expr = brm(
    new_cases ~ 
      I( (age-62) / 30) * consent +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent +
      sex_female * consent,
    family = bernoulli(link = "logit"),
    data = data_ckd,
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = priors,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_complex_lin_NewCases",
  reuse = TRUE
)

prior_summary(model_complex_lin_NewCases)
summary(model_complex_lin_NewCases, robust = TRUE)

exp(fixef(model_complex_lin_NewCases, robust = TRUE))[,c(1, 3, 4)]
```

#### Extract posterior

```{r}
consent_prop <- mean(data_ckd$consent)

age_scale <- 25 / 30

post <- posterior::as_draws_df(
  model_complex_lin_NewCases
)

b_all <- post %>%
  dplyr::select(
    dplyr::starts_with("b_")
  )

base <- b_all %>%
  dplyr::transmute(
    Intercept = b_Intercept,
    age_25y = b_IageM62D30 * age_scale,
    sex_female = b_sex_female,
    transplant = b_transplant,
    FU_DM1 = b_FU_DM1,
    FU_DM2 = b_FU_DM2,
    FU_steatosis = b_FU_steatosis,
    FU_hypertension = b_FU_hypertension,
    FU_atherosclerosis = b_FU_atherosclerosis,
    HF_or_AF = b_HF_or_AF
  )

eff_consent0 <- base

eff_consent1 <- base %>%
  dplyr::mutate(
    Intercept = Intercept + b_all$b_consent,
    age_25y = age_25y + b_all$`b_IageM62D30:consent` * age_scale,
    transplant = transplant + b_all$`b_consent:transplant`,
    FU_DM1 = FU_DM1 + b_all$`b_consent:FU_DM1`,
    FU_DM2 = FU_DM2 + b_all$`b_consent:FU_DM2`,
    FU_steatosis = FU_steatosis + b_all$`b_consent:FU_steatosis`,
    FU_hypertension = FU_hypertension + b_all$`b_consent:FU_hypertension`,
    FU_atherosclerosis = FU_atherosclerosis + b_all$`b_consent:FU_atherosclerosis`,
    HF_or_AF = HF_or_AF + b_all$`b_consent:HF_or_AF`,
    sex_female = sex_female + b_all$`b_consent:sex_female`
  )

eff_consent_mean <- base %>%
  dplyr::mutate(
    Intercept = Intercept + b_all$b_consent * consent_prop,
    age_25y = age_25y + b_all$`b_IageM62D30:consent` * consent_prop * age_scale,
    transplant = transplant + b_all$`b_consent:transplant` * consent_prop,
    FU_DM1 = FU_DM1 + b_all$`b_consent:FU_DM1` * consent_prop,
    FU_DM2 = FU_DM2 + b_all$`b_consent:FU_DM2` * consent_prop,
    FU_steatosis = FU_steatosis + b_all$`b_consent:FU_steatosis` * consent_prop,
    FU_hypertension = FU_hypertension + b_all$`b_consent:FU_hypertension` * consent_prop,
    FU_atherosclerosis = FU_atherosclerosis + b_all$`b_consent:FU_atherosclerosis` * consent_prop,
    HF_or_AF = HF_or_AF + b_all$`b_consent:HF_or_AF` * consent_prop,
    sex_female = sex_female + b_all$`b_consent:sex_female` * consent_prop
  )


data_dm_ref <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 0
  )

p0_dm <- brms::posterior_epred(
  model_complex_lin_NewCases,
  newdata = data_dm_ref
) %>%
  rowMeans()

odds0_dm <- p0_dm / (1 - p0_dm)

data_dm1 <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 1,
    FU_DM2 = 0
  )

p1_dm1 <- brms::posterior_epred(
  model_complex_lin_NewCases,
  newdata = data_dm1
) %>%
  rowMeans()

odds1_dm1 <- p1_dm1 / (1 - p1_dm1)

data_dm2 <- data_ckd %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 1
  )

p1_dm2 <- brms::posterior_epred(
  model_complex_lin_NewCases,
  newdata = data_dm2
) %>%
  rowMeans()

odds1_dm2 <- p1_dm2 / (1 - p1_dm2)

log_or_dm1_marg <- log(odds1_dm1 / odds0_dm)
log_or_dm2_marg <- log(odds1_dm2 / odds0_dm)

eff_marginal <- dplyr::tibble(
  age_25y = calc_marg_log_or(
    model_complex_lin_NewCases,
    data_ckd,
    "age",
    75,
    50
  ),
  sex_female = calc_marg_log_or(
    model_complex_lin_NewCases,
    data_ckd,
    "sex_female",
    1,
    0
  ),
  transplant = calc_marg_log_or(
    model_complex_lin_NewCases,
    data_ckd,
    "transplant",
    1,
    0
  ),
  FU_DM1 = log_or_dm1_marg,
  FU_DM2 = log_or_dm2_marg,
  FU_steatosis = calc_marg_log_or(
    model_complex_lin_NewCases,
    data_ckd,
    "FU_steatosis",
    1,
    0
  ),
  FU_hypertension = calc_marg_log_or(
    model_complex_lin_NewCases,
    data_ckd,
    "FU_hypertension",
    1,
    0
  ),
  FU_atherosclerosis = calc_marg_log_or(
    model_complex_lin_NewCases,
    data_ckd,
    "FU_atherosclerosis",
    1,
    0
  ),
  HF_or_AF = calc_marg_log_or(
    model_complex_lin_NewCases,
    data_ckd,
    "HF_or_AF",
    1,
    0
  )
)

summ <- dplyr::bind_rows(
  posterior::summarise_draws(
    eff_consent0,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "unscreened"
    ),
  posterior::summarise_draws(
    eff_consent1,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "screening"
    ),
  posterior::summarise_draws(
    eff_consent_mean,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "enriched_std_cond"
    ),
  posterior::summarise_draws(
    eff_marginal,
    median,
    ~ quantile(.x, c(0.025, 0.975))
  ) %>%
    dplyr::mutate(
      cohort = "enriched_marginal"
    )
) %>%
  dplyr::rename(
    predictor = variable,
    CI_L = `2.5%`,
    CI_U = `97.5%`
  ) %>%
  dplyr::mutate(
    OR = exp(median),
    CI_L = exp(CI_L),
    CI_U = exp(CI_U)
  )

data_forest <- summ %>%
  dplyr::filter(
    predictor != "Intercept"
  ) %>%
  dplyr::mutate(
    predictor = factor(
      predictor,
      levels = rev(
        unique(
          predictor
        )
      )
    )
  )

tab_ready <- summ %>%
  dplyr::filter(
    predictor != "Intercept"
  ) %>%
  dplyr::mutate(
    label = sprintf(
      "%.2f [%.2f, %.2f]",
      OR,
      CI_L,
      CI_U
    )
  ) %>%
  dplyr::select(
    cohort,
    predictor,
    label
  )

tab_wide <- tab_ready %>%
  tidyr::pivot_wider(
    names_from = cohort,
    values_from = label
  ) %>%
  dplyr::arrange(
    predictor
  )

kableExtra::kable(tab_wide, caption = "Associations between clinical factors and newly diagnosed CKD from the multivariable Bayesian logistic regression model. Odds ratios (ORs) are shown with 95% credible intervals. For screened and unscreened patients, ORs are conditional estimates from the interaction model. For the Enriched cohort, two summary measures are presented: the population-average conditional OR, calculated as the weighted average of screened- and unscreened-specific log-ORs using their observed proportions in the cohort, and the marginal OR, obtained by posterior standardisation of predicted probabilities across all patients in the Enriched cohort. Age effects are expressed for a 25-year difference; the marginal age OR compares 75 versus 50 years.")

```

#### Visualisation, screening + enriched conditional

```{r, fig.height=5, fig.width=6.5}

data_forest %>%
  dplyr::filter(
    cohort != "unscreened",
    cohort != "enriched_marginal"
  ) %>%
  dplyr::mutate(
    predictor = dplyr::recode(
      predictor,
      "age_25y" = "Age (per 25y)",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_hypertension" = "Arterial hypertension",
      "HF_or_AF" = "Heart failure or atrial fibrillation",
      "FU_atherosclerosis" = "ASCVD",
      "FU_steatosis" = "Liver steatosis",
      "sex_female" = "Female"
    ),
    predictor = factor(
      predictor,
      levels = rev(
        unique(
          predictor
        )
      )
    ),
    cohort = factor(
      cohort,
      levels = c(
        "unscreened",
        "screening",
        "enriched_std_cond",
        "enriched_marginal"
      ),
      labels = c(
        "Unscreened patients",
        "Screening cohort",
        "Enriched cohort (conditional)",
        "Enriched cohort (marginal)"
      )
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = OR,
      y = predictor,
      color = cohort
    )
  ) +
  ggplot2::geom_point(
    position = ggplot2::position_dodge(width = 0.6),
    size = 2
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(
      xmin = CI_L,
      xmax = CI_U
    ),
    position = ggplot2::position_dodge(width = 0.6),
    height = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "gray40"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  ggplot2::scale_color_manual(
    values = c(
      "Unscreened patients" = "#1b9e77",
      "Screening cohort" = "#d95f02",
      "Enriched cohort (conditional)" = "skyblue3",
      "Enriched cohort (marginal)" = "#984ea3"
    )
  ) +
  ggplot2::labs(
    x = "Adjusted odds ratio (log scale) for newly diagnosed CKD",
    y = NULL,
    color = NULL
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      nrow = 1
    )
  ) +
  ggplot2::theme_minimal(
    base_size = 12
  ) +
  ggplot2::theme(
    legend.position = "top",
    axis.text.y = ggplot2::element_text(size = 11),
    legend.text = ggplot2::element_text(size = 10)
  )

```

#### Visualisation, all

```{r, fig.height=5, fig.width=6.5}

plotac <- "suppl_fig_5"
path <- "gitignore/figures"

fig <- data_forest %>%
  dplyr::mutate(
    predictor = dplyr::recode(
      predictor,
      "age_25y" = "Age (per 25y)",
      "transplant" = "Non-renal transplantation",
      "FU_DM1" = "T1DM",
      "FU_DM2" = "T2DM",
      "FU_hypertension" = "Arterial hypertension",
      "HF_or_AF" = "Heart failure or atrial fibrillation",
      "FU_atherosclerosis" = "ASCVD",
      "FU_steatosis" = "Liver steatosis",
      "sex_female" = "Female"
    ),
    predictor = factor(
      predictor,
      levels = rev(
        unique(
          predictor
        )
      )
    ),
    cohort = factor(
      cohort,
      levels = c(
        "unscreened",
        "screening",
        "enriched_std_cond",
        "enriched_marginal"
      ),
      labels = c(
        "Unscreened patients",
        "Screening cohort",
        "Enriched cohort (conditional)",
        "Enriched cohort (marginal)"
      )
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = OR,
      y = predictor,
      color = cohort
    )
  ) +
  ggplot2::geom_point(
    position = ggplot2::position_dodge(width = 0.6),
    size = 2
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(
      xmin = CI_L,
      xmax = CI_U
    ),
    position = ggplot2::position_dodge(width = 0.6),
    height = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "gray40"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(1 / 4, 1 / 2, 1, 2, 4, 8),
    labels = c("1/4", "1/2", "1", "2", "4", "8")
  ) +
  ggplot2::scale_color_manual(
    values = c(
      "Unscreened patients" = "#1b9e77",
      "Screening cohort" = "#d95f02",
      "Enriched cohort (conditional)" = "skyblue3",
      "Enriched cohort (marginal)" = "#984ea3"
    )
  ) +
  ggplot2::labs(
    x = "Adjusted odds ratio (log scale) for newly diagnosed CKD",
    y = NULL,
    color = NULL
  ) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      nrow = 2
    )
  ) +
  ggplot2::theme_minimal(
    base_size = 12
  ) +
  ggplot2::theme(
    legend.position = "top",
    axis.text.y = ggplot2::element_text(size = 11),
    legend.text = ggplot2::element_text(size = 10)
  )

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggplot2::ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 6.5,
    height = 5
  )
}

```

#### Sensitivity analysis

Fit model

```{r}
model_sens_lin_NewCases <- run(
  expr = brm(
    new_cases ~ 
      s(I( (age-62) / 30), by = consent, k = 5) +
      transplant * consent +
      FU_DM1 * consent +
      FU_DM2 * consent +
      sex_female * consent + 
      FU_steatosis * consent +
      FU_hypertension * consent +
      FU_atherosclerosis * consent +
      HF_or_AF * consent +
      BMI_over30 * consent,
    data = data_ckd |> 
      dplyr::filter(!is.na(BMI_over30)) |>
      dplyr::mutate(consent = factor(consent)),
    family = bernoulli(link = "logit"),
    backend = "cmdstanr",
    seed = 736,
    cores = 1, chains = 2,
    prior = priors_sens,
    iter = 7000, warmup = 2000,
    control = list(adapt_delta = 0.95)
  ),
  path = "gitignore/run/model_sens_lin_NewCases",
  reuse = TRUE
)

prior_summary(model_sens_lin_NewCases )
summary(model_sens_lin_NewCases , robust = TRUE)

exp(fixef(model_sens_lin_NewCases , robust = TRUE))[,c(1, 3, 4)]
```

Extract posteriors and show results

```{r}
dummy_base <- data_ckd_complete[1, ] %>%
  dplyr::mutate(
    transplant = 0,
    FU_DM1 = 0,
    FU_DM2 = 0,
    sex_female = 0,
    FU_steatosis = 0,
    FU_hypertension = 0,
    FU_atherosclerosis = 0,
    HF_or_AF = 0,
    BMI_over30 = 0
  )

d_c0_a50 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(0, levels = c("0", "1")),
    age = 50
  )

d_c0_a75 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(0, levels = c("0", "1")),
    age = 75
  )

lin_c0_a50 <- posterior_linpred(
  model_sens_lin_NewCases,
  newdata = d_c0_a50
)

lin_c0_a75 <- posterior_linpred(
  model_sens_lin_NewCases,
  newdata = d_c0_a75
)

age_25y_c0 <- as.numeric(lin_c0_a75 - lin_c0_a50)

d_c1_a50 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(1, levels = c("0", "1")),
    age = 50
  )

d_c1_a75 <- dummy_base %>%
  dplyr::mutate(
    consent = factor(1, levels = c("0", "1")),
    age = 75
  )

lin_c1_a50 <- posterior_linpred(
  model_sens_lin_NewCases,
  newdata = d_c1_a50
)

lin_c1_a75 <- posterior_linpred(
  model_sens_lin_NewCases,
  newdata = d_c1_a75
)

age_25y_c1 <- as.numeric(lin_c1_a75 - lin_c1_a50)

post <- as_draws_df(
  model_sens_lin_NewCases
)

b_all <- post %>%
  dplyr::select(
    dplyr::starts_with("b_")
  )

eff_consent0 <- dplyr::tibble(
  age_25y = age_25y_c0,
  sex_female = b_all$b_sex_female,
  transplant = b_all$b_transplant,
  FU_DM1 = b_all$b_FU_DM1,
  FU_DM2 = b_all$b_FU_DM2,
  FU_steatosis = b_all$b_FU_steatosis,
  FU_hypertension = b_all$b_FU_hypertension,
  FU_atherosclerosis = b_all$b_FU_atherosclerosis,
  HF_or_AF = b_all$b_HF_or_AF,
  BMI_over30 = b_all$b_BMI_over30
)

eff_consent1 <- dplyr::tibble(
  age_25y = age_25y_c1,
  sex_female = b_all$b_sex_female + b_all$`b_consent1:sex_female`,
  transplant = b_all$b_transplant + b_all$`b_transplant:consent1`,
  FU_DM1 = b_all$b_FU_DM1 + b_all$`b_consent1:FU_DM1`,
  FU_DM2 = b_all$b_FU_DM2 + b_all$`b_consent1:FU_DM2`,
  FU_steatosis = b_all$b_FU_steatosis + b_all$`b_consent1:FU_steatosis`,
  FU_hypertension = b_all$b_FU_hypertension + b_all$`b_consent1:FU_hypertension`,
  FU_atherosclerosis = b_all$b_FU_atherosclerosis + b_all$`b_consent1:FU_atherosclerosis`,
  HF_or_AF = b_all$b_HF_or_AF + b_all$`b_consent1:HF_or_AF`,
  BMI_over30 = b_all$b_BMI_over30 + b_all$`b_consent1:BMI_over30`
)

eff_consent_mean <- dplyr::tibble(
  age_25y = age_25y_c0 * (1 - consent_prop) + 
    age_25y_c1 * consent_prop,
  sex_female = eff_consent0$sex_female * (1 - consent_prop) + 
    eff_consent1$sex_female * consent_prop,
  transplant = eff_consent0$transplant * (1 - consent_prop) + 
    eff_consent1$transplant * consent_prop,
  FU_DM1 = eff_consent0$FU_DM1 * (1 - consent_prop) + 
    eff_consent1$FU_DM1 * consent_prop,
  FU_DM2 = eff_consent0$FU_DM2 * (1 - consent_prop) + 
    eff_consent1$FU_DM2 * consent_prop,
  FU_steatosis = eff_consent0$FU_steatosis * (1 - consent_prop) + 
    eff_consent1$FU_steatosis * consent_prop,
  FU_hypertension = eff_consent0$FU_hypertension * (1 - consent_prop) + 
    eff_consent1$FU_hypertension * consent_prop,
  FU_atherosclerosis = eff_consent0$FU_atherosclerosis * (1 - consent_prop) + 
    eff_consent1$FU_atherosclerosis * consent_prop,
  HF_or_AF = eff_consent0$HF_or_AF * (1 - consent_prop) + 
    eff_consent1$HF_or_AF * consent_prop,
  BMI_over30 = eff_consent0$BMI_over30 * (1 - consent_prop) + 
    eff_consent1$BMI_over30 * consent_prop
)

calc_marg_log_or <- function(model, data, var_name, val1, val0) {
  d1 <- data
  
  d1[[var_name]] <- val1
  
  p1 <- posterior_epred(
    model,
    newdata = d1
  )
  
  p1_marg <- rowMeans(p1)

  d0 <- data
  
  d0[[var_name]] <- val0
  
  p0 <- posterior_epred(
    model,
    newdata = d0
  )
  
  p0_marg <- rowMeans(p0)

  odds1 <- p1_marg / (1 - p1_marg)
  
  odds0 <- p0_marg / (1 - p0_marg)

  marg_or <- odds1 / odds0
  
  return(log(marg_or))
}

data_dm_ref <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 0
  )

p0_dm <- posterior_epred(
  model_sens_lin_NewCases,
  newdata = data_dm_ref
) %>%
  rowMeans()

odds0_dm <- p0_dm / (1 - p0_dm)

data_dm1 <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 1,
    FU_DM2 = 0
  )

p1_dm1 <- posterior_epred(
  model_sens_lin_NewCases,
  newdata = data_dm1
) %>%
  rowMeans()

odds1_dm1 <- p1_dm1 / (1 - p1_dm1)

data_dm2 <- data_ckd_complete %>%
  dplyr::mutate(
    FU_DM1 = 0,
    FU_DM2 = 1
  )

p1_dm2 <- posterior_epred(
  model_sens_lin_NewCases,
  newdata = data_dm2
) %>%
  rowMeans()

odds1_dm2 <- p1_dm2 / (1 - p1_dm2)

log_or_dm1_marg <- log(odds1_dm1 / odds0_dm)

log_or_dm2_marg <- log(odds1_dm2 / odds0_dm)

eff_marginal <- dplyr::tibble(
  age_25y = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "age",
    75,
    50
  ),
  sex_female = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "sex_female",
    1,
    0
  ),
  transplant = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "transplant",
    1,
    0
  ),
  FU_DM1 = log_or_dm1_marg,
  FU_DM2 = log_or_dm2_marg,
  FU_steatosis = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "FU_steatosis",
    1,
    0
  ),
  FU_hypertension = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "FU_hypertension",
    1,
    0
  ),
  FU_atherosclerosis = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "FU_atherosclerosis",
    1,
    0
  ),
  HF_or_AF = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "HF_or_AF",
    1,
    0
  ),
  BMI_over30 = calc_marg_log_or(
    model_sens_lin_NewCases,
    data_ckd_complete,
    "BMI_over30",
    1,
    0
  )
)

summ <- dplyr::bind_rows(
  summarise_draws(
    eff_consent0,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "unscreened"
    ),
  summarise_draws(
    eff_consent1,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "screening"
    ),
  summarise_draws(
    eff_consent_mean,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_std_cond"
    ),
  summarise_draws(
    eff_marginal,
    median,
    ~ quantile(
      .x,
      c(0.025, 0.975)
    )
  ) %>%
    dplyr::mutate(
      cohort = "enriched_marginal"
    )
) %>%
  dplyr::rename(
    predictor = variable,
    CI_L = `2.5%`,
    CI_U = `97.5%`
  ) %>%
  dplyr::mutate(
    OR = exp(median),
    CI_L = exp(CI_L),
    CI_U = exp(CI_U)
  )

tab_ready <- summ %>%
  dplyr::mutate(
    label = sprintf(
      "%.2f [%.2f, %.2f]",
      OR,
      CI_L,
      CI_U
    )
  ) %>%
  dplyr::select(
    cohort,
    predictor,
    label
  )

tab_wide <- tab_ready %>%
  tidyr::pivot_wider(
    names_from = cohort,
    values_from = label
  ) %>%
  dplyr::arrange(
    predictor
  )

kableExtra::kable(
  tab_wide,
  caption = "Associations between clinical factors and newly diagnosed CKD cases from the multivariable Bayesian logistic regression model. Odds ratios (ORs) are shown with 95% credible intervals. For screened and unscreened patients, ORs are conditional estimates from the interaction model. For the Enriched cohort, two summary measures are presented: the population-average conditional OR, calculated as the weighted average of screened- and unscreened-specific log-ORs using their observed proportions in the cohort, and the marginal OR, obtained by posterior standardisation of predicted probabilities across all patients in the Enriched cohort. Age effects are expressed for a 25-year difference comparing 75 versus 50 years."
)
```


# Model-projected gain in dialysis-free time

## Propotion of patients taking given drugs from those egligable

```{r}

## RAASi
data_ckd %>% filter(FU_RAASi == 1, elig_RAASi_all == 1) %>% nrow() / 
  data_ckd %>% filter(elig_RAASi_all == 1) %>% nrow()

## SGLT2
data_ckd %>% filter(FU_SGLT2 == 1, elig_SGLT2i_total == 1) %>% nrow() / 
  data_ckd %>% filter(elig_SGLT2i_total == 1) %>% nrow()

## Finerenone
data_ckd %>% filter(FU_finerenone == 1, elig_finerenone == 1) %>% nrow() / 
  data_ckd %>% filter(elig_finerenone == 1) %>% nrow()
```

## Who are the eligible CKD patients not receiving treatment?

### RAASi

```{r}
data_ckd %>% 
  dplyr::filter(
    elig_RAASi_all == 1, 
    identified_CKD_by_lab == 1
    ) %>% 
  dplyr::mutate(FU_RAASi = if_else(FU_RAASi == 1, 'user', 'non-user')) %>%  
  dplyr::select(age, consent, sex, transplant:FU_finerenone,
                uACR_value:eGFR_text, CKD_record:new_cases,
                elig_RAASi_all:elig_finerenone
                ) %>% 
  gtsummary::tbl_summary(by = FU_RAASi) %>% 
  add_p()
```

### SGLT2

```{r}
data_ckd %>% 
  dplyr::filter(
    elig_SGLT2i_total == 1, 
    identified_CKD_by_lab == 1
    ) %>% 
  dplyr::mutate(FU_SGLT2 = if_else(FU_SGLT2 == 1, 'user', 'non-user')) %>%  
  dplyr::select(age, consent, sex, transplant:FU_finerenone,
                uACR_value:eGFR_text, CKD_record:new_cases,
                elig_RAASi_all:elig_finerenone
                ) %>% 
  gtsummary::tbl_summary(by = FU_SGLT2) %>% 
  add_p()
```

### GLP1_RA

```{r}
data_ckd %>% 
  dplyr::filter(
    elig_GLP1_RA == 1, 
    identified_CKD_by_lab == 1
    ) %>% 
  dplyr::mutate(FU_GLP1 = if_else(FU_GLP1 == 1, 'user', 'non-user')) %>%  
  dplyr::select(age, consent, sex, transplant:FU_finerenone,
                uACR_value:eGFR_text, CKD_record:new_cases,
                elig_RAASi_all:elig_finerenone
                ) %>% 
  gtsummary::tbl_summary(by = FU_GLP1) %>% 
  add_p()
```

### Finerenone

```{r}
data_ckd %>% 
  dplyr::filter(
    elig_finerenone == 1, 
    identified_CKD_by_lab == 1
    ) %>% 
  dplyr::mutate(FU_finerenone = if_else(FU_finerenone == 1, 'user', 'non-user')) %>%  
  dplyr::select(age, consent, sex, transplant:FU_finerenone,
                uACR_value:eGFR_text, CKD_record:new_cases,
                elig_RAASi_all:elig_finerenone
                ) %>% 
  gtsummary::tbl_summary(by = FU_finerenone) %>% 
  add_p()
```

## EGFR slopes across treatments

```{r}
egfr_slope_plcb <- -3.59
egfr_slope_GLP1 <- egfr_slope_plcb + 1.16
egfr_slope_sglt2 <- -1.67
egfr_slope_finerenone <- -2.66
egfr_slope_sglt2_finerenone <- egfr_slope_plcb - 
  (egfr_slope_plcb - egfr_slope_sglt2) - 
  (egfr_slope_plcb - egfr_slope_finerenone)

egfr_slope_plcb
egfr_slope_GLP1
egfr_slope_sglt2
egfr_slope_finerenone
egfr_slope_sglt2_finerenone

sglt2_add <-  egfr_slope_sglt2 - egfr_slope_plcb 
finerenone_add <- egfr_slope_finerenone - egfr_slope_plcb 
GLP1_add <- egfr_slope_GLP1 - egfr_slope_plcb

sglt2_add
finerenone_add
GLP1_add
```

## Survival

Impute data with a single stochastic imputation

```{r}
if (!file.exists('data/data_survival_imp.rds')) {
  vars_no_impute <- c(
    "EntryDate", "Sex", "time", "event"
  )

  bin_vars <- c(
    "FU_DM", "FU_hypertension", "FU_steatosis",
    "FU_atrial_fibrillation", "FU_heart_failure",
    "ischemic_heart_disease", "FU_atherosclerosis_PAD",
    "FU_stroke"
  )

  data_survival[bin_vars] <- lapply(
    data_survival[bin_vars],
    function(x) factor(x, levels = c(0, 1))
  )

  meth <- make.method(data_survival)
  pred <- make.predictorMatrix(data_survival)

  meth[vars_no_impute] <- ""
  pred[vars_no_impute, ] <- 0
  pred[, vars_no_impute] <- 0

  meth[bin_vars] <- "logreg"

  imp <- mice(
    data_survival,
    m = 1,
    method = meth,
    predictorMatrix = pred,
    print = FALSE,
    seed = seeding
  )

  data_survival_imp <- complete(imp, 1)

  data_survival_imp[bin_vars] <- lapply(
    data_survival_imp[bin_vars],
    function(x) as.numeric(as.character(x))
  )

  saveRDS(data_survival_imp, "data/data_survival_imp.rds")
}

data_survival <- readRDS("data/data_survival_imp.rds")
```

### Fit model

```{r}

model_surv <- coxph(
  Surv(time, event) ~ 
    transplant + 
    FU_DM + 
    FU_hypertension +
    sex_female +
    HF_or_AF + 
    CKD_record + 
    FU_atherosclerosis + 
    FU_steatosis + 
    pspline(age, df = 5),
  data = data_survival
)

anova(model_surv)
sf <- survfit(model_surv, newdata = data_ckd)

pred_cox <- sapply(
  1:ncol(sf$surv),
  function(i) {
    t <- sf$time
    s <- sf$surv[, i]
    sum(diff(t) * (head(s, -1) + tail(s, -1)) / 2)
  }
)

summary(pred_cox)
```

### Visualise cox

```{r}
newdat <- tibble::tibble(
  age = c(20, 80, 20, 80),
  transplant = c(0, 0, 1, 1),
  FU_DM = 0,
  FU_hypertension = 0,
  sex_female = 0,
  HF_or_AF = 0,
  CKD_record = 1,
  FU_atherosclerosis = 0,
  FU_steatosis = 0
)

sf <- survfit(model_surv, newdata = newdat)

plot(sf, col = c("skyblue1", "seagreen3", "pink2", "hotpink2"), lwd = 2,
     xlab = "time", ylab = "survival")
legend(
  "bottomleft",
  legend = c("age 20 noTX", "age 80 no TX", "age 20 TX", "age 80 TX"),
  col = c("skyblue1", "seagreen3", "pink2", "hotpink2"),
  lwd = 2
)
```

### Function to get dialysis-free survival prediction

```{r}



sf_patient <- survival::survfit(
  model_surv,
  newdata = data_ckd
)

surv_horizon <- max(
  sf_patient$time,
  na.rm = TRUE
)

surv_horizon

stopifnot(
  ncol(sf_patient$surv) == nrow(data_ckd)
)



calc_expected_dialysis_free <- function(sf_object, time_to_RF, horizon) {
  
  tau_all <- pmin(pmax(time_to_RF, 0), horizon)
  
  purrr::map_dbl(seq_along(tau_all), function(i) {
    
    tau <- tau_all[i]
    
    if (is.na(tau)) {
      return(NA_real_)
    }
    
    idx <- sf_object$time < tau
    
    time_points <- c(0, sf_object$time[idx], tau)
    survival_prob <- c(1, sf_object$surv[idx, i])
    
    sum(diff(time_points) * survival_prob)
  })
}


```

## Individual eGFR slopes and individual years on dialysis saved

```{r}
smallest_possible_egfr <- -0.5


data_ckd <- data_ckd %>%
  dplyr::mutate(
    eGFR_min_to10 = dplyr::if_else(
      eGFR_min > 10,
      eGFR_min - 10,
      0
    )
  ) %>%
  dplyr::mutate(
    egfr_slope = pmin(
      smallest_possible_egfr,
      egfr_slope_plcb +
        sglt2_add * FU_SGLT2 +
        finerenone_add * FU_finerenone +
        GLP1_add * FU_GLP1
    ),
    egfr_slope_potential = pmin(
      smallest_possible_egfr,
      egfr_slope_plcb +
        sglt2_add * pmax(FU_SGLT2, elig_SGLT2i_total) +
        finerenone_add * pmax(FU_finerenone, elig_finerenone) +
        GLP1_add * pmax(FU_GLP1, elig_GLP1_RA)
    ),
    egfr_slope_pot_SGLT = pmin(
      smallest_possible_egfr,
      egfr_slope_plcb +
        sglt2_add * pmax(FU_SGLT2, elig_SGLT2i_total) +
        finerenone_add * FU_finerenone +
        GLP1_add * FU_GLP1
    ),
    egfr_slope_pot_GLP = pmin(
      smallest_possible_egfr,
      egfr_slope_plcb +
        sglt2_add * FU_SGLT2 +
        finerenone_add * FU_finerenone +
        GLP1_add * pmax(FU_GLP1, elig_GLP1_RA)
    ),
    egfr_slope_pot_finerenone = pmin(
      smallest_possible_egfr,
      egfr_slope_plcb +
        sglt2_add * FU_SGLT2 +
        finerenone_add * pmax(FU_finerenone, elig_finerenone) +
        GLP1_add * FU_GLP1
    )
  ) %>%
  dplyr::mutate(
    new_SGLT = pmax(elig_SGLT2i_total - FU_SGLT2, 0),
    new_GLP = pmax(elig_GLP1_RA - FU_GLP1, 0),
    new_Fin = pmax(elig_finerenone - FU_finerenone, 0),
    add_new_SGLT = sglt2_add * new_SGLT,
    add_new_GLP = GLP1_add * new_GLP,
    add_new_Fin = finerenone_add * new_Fin,
    max_new_eff = pmax(add_new_SGLT, add_new_GLP, add_new_Fin),
    egfr_slope_partialAdd = pmin(
      smallest_possible_egfr,
      egfr_slope +
        if_else(
          add_new_SGLT == max_new_eff,
          add_new_SGLT,
          0.5 * add_new_SGLT
        ) +
        if_else(
          add_new_GLP == max_new_eff,
          add_new_GLP,
          0.5 * add_new_GLP
        ) +
        if_else(
          add_new_Fin == max_new_eff,
          add_new_Fin,
          0.5 * add_new_Fin
        )
    )
  ) %>%
  dplyr::mutate(
    years_to_RF_curr =
      eGFR_min_to10 / (-egfr_slope),
    years_to_RF_pot =
      eGFR_min_to10 / (-egfr_slope_potential),
    years_to_RF_partial =
      eGFR_min_to10 / (-egfr_slope_partialAdd),
    years_to_RF_pot_SGLT =
      eGFR_min_to10 / (-egfr_slope_pot_SGLT),
    years_to_RF_pot_GLP =
      eGFR_min_to10 / (-egfr_slope_pot_GLP),
    years_to_RF_pot_fin =
      eGFR_min_to10 / (-egfr_slope_pot_finerenone)
  ) %>%
  dplyr::mutate(
    years_to_RF_curr =
      pmax(years_to_RF_curr, 0),
    years_to_RF_pot =
      pmax(years_to_RF_pot, 0),
    years_to_RF_partial =
      pmax(years_to_RF_partial, 0),
    years_to_RF_pot_SGLT =
      pmax(years_to_RF_pot_SGLT, 0),
    years_to_RF_pot_GLP =
      pmax(years_to_RF_pot_GLP, 0),
    years_to_RF_pot_fin =
      pmax(years_to_RF_pot_fin, 0)
  ) %>%
  dplyr::mutate(
    DialysisMonthsSaved_byTreat =
      (years_to_RF_pot - years_to_RF_curr) * 12,
    DialysisMonthsSaved_byTreatPartial =
      (years_to_RF_partial - years_to_RF_curr) * 12,
    DialysisMonthsSaved_bySGLT =
      (years_to_RF_pot_SGLT - years_to_RF_curr) * 12,
    DialysisMonthsSaved_byGLP =
      (years_to_RF_pot_GLP - years_to_RF_curr) * 12,
    DialysisMonthsSaved_byFin =
      (years_to_RF_pot_fin - years_to_RF_curr) * 12
  )

data_ckd <- data_ckd %>%
  dplyr::mutate(
    years_to_RF_curr_SurvCorr =
      calc_expected_dialysis_free(
        sf_patient,
        years_to_RF_curr,
        surv_horizon
      ),
    years_to_RF_pot_SurvCorr =
      calc_expected_dialysis_free(
        sf_patient,
        years_to_RF_pot,
        surv_horizon
      ),
    years_to_RF_partial_SurvCorr =
      calc_expected_dialysis_free(
        sf_patient,
        years_to_RF_partial,
        surv_horizon
      ),
    years_to_RF_pot_SGLT_SurvCorr =
      calc_expected_dialysis_free(
        sf_patient,
        years_to_RF_pot_SGLT,
        surv_horizon
      ),
    years_to_RF_pot_GLP_SurvCorr =
      calc_expected_dialysis_free(
        sf_patient,
        years_to_RF_pot_GLP,
        surv_horizon
      ),
    years_to_RF_pot_fin_SurvCorr =
      calc_expected_dialysis_free(
        sf_patient,
        years_to_RF_pot_fin,
        surv_horizon
      )
  ) %>%
  dplyr::mutate(
    DialysisMonthsSaved_byTreat_SurvCorr =
      (years_to_RF_pot_SurvCorr -
         years_to_RF_curr_SurvCorr) * 12,
    DialysisMonthsSaved_byTreatPartial_SurvCorr =
      (years_to_RF_partial_SurvCorr -
         years_to_RF_curr_SurvCorr) * 12,
    DialysisMonthsSaved_bySGLT_SurvCorr =
      (years_to_RF_pot_SGLT_SurvCorr -
         years_to_RF_curr_SurvCorr) * 12,
    DialysisMonthsSaved_byGLP_SurvCorr =
      (years_to_RF_pot_GLP_SurvCorr -
         years_to_RF_curr_SurvCorr) * 12,
    DialysisMonthsSaved_byFin_SurvCorr =
      (years_to_RF_pot_fin_SurvCorr -
         years_to_RF_curr_SurvCorr) * 12
  )

summary(
  data_ckd %>%
    dplyr::select(
      years_to_RF_curr_SurvCorr,
      years_to_RF_pot_SurvCorr,
      years_to_RF_partial_SurvCorr,
      DialysisMonthsSaved_byTreat_SurvCorr,
      DialysisMonthsSaved_byTreatPartial_SurvCorr,
      DialysisMonthsSaved_bySGLT_SurvCorr,
      DialysisMonthsSaved_byGLP_SurvCorr,
      DialysisMonthsSaved_byFin_SurvCorr
    )
)

```

## Tabular results

### All CKD population

```{r}
data_ckd %>%
    filter(identified_CKD_by_lab == 1) %>%
    select(
        sex, age,
        FU_RAASi, FU_SGLT2, FU_finerenone, FU_GLP1,
        elig_RAASi_all, elig_SGLT2i_total, elig_finerenone, elig_GLP1_RA,
        DialysisMonthsSaved_byTreat_SurvCorr,
        DialysisMonthsSaved_byTreatPartial_SurvCorr,
        DialysisMonthsSaved_bySGLT_SurvCorr,
        DialysisMonthsSaved_byGLP_SurvCorr,
        DialysisMonthsSaved_byFin_SurvCorr
    ) %>%
    mutate(
        across(
            where(~ all(na.omit(.) %in% c(0, 1))),
            ~ factor(., levels = c(0, 1))
        )
    ) %>%
    tbl_summary(
        type = list(
            everything() ~ "continuous",
            where(is.factor) ~ "categorical"
        ),
        statistic = list(
            all_continuous() ~ "{mean}, [{min}, {max}]",
            all_categorical() ~ "{n} ({p}%)"
        ),
        digits = all_continuous() ~ c(2, 2, 2)
    ) %>%
    modify_caption(
        '*CKD population identified by laboratory measurements in IKEM. Includes
    treatment used ("FU_[...]"), treatment eligibility ("elig_[...]"), and
    average [range] dialysis-free months with or without survival correction.*'
    )
```

### SGLT2-egligable

```{r}
data_ckd %>% dplyr::filter(
  identified_CKD_by_lab == 1, elig_SGLT2i_total == 1
) %>% select(
  sex, age,
  FU_RAASi, FU_SGLT2, FU_finerenone, FU_GLP1,
  elig_RAASi_all, elig_SGLT2i_total, elig_finerenone, elig_GLP1_RA,
  DialysisMonthsSaved_byTreat_SurvCorr,
  DialysisMonthsSaved_byTreatPartial_SurvCorr,
  DialysisMonthsSaved_bySGLT_SurvCorr
  ) %>%
    mutate(
        across(
            where(~ all(na.omit(.) %in% c(0, 1))),
            ~ factor(., levels = c(0, 1))
        )
    ) %>%
    tbl_summary(
        type = list(
            everything() ~ "continuous",
            where(is.factor) ~ "categorical"
        ),
        statistic = list(
            all_continuous() ~ "{mean}, [{min}, {max}]",
            all_categorical() ~ "{n} ({p}%)"
        ),
        digits = all_continuous() ~ c(2, 2, 2)
    ) %>%
  modify_caption('*CKD population **egigable for SGLT2i**. \
Includes treatment used (\"FU_[...]\"), treatment eligibility (\"elig_[...]\"), \
and average [range] dialysis-free months with or without survival correction.*')
```

### Finerone-egligable

```{r}
data_ckd %>% dplyr::filter(
  identified_CKD_by_lab == 1, elig_finerenone == 1
) %>% select(
  sex, age,
  FU_RAASi, FU_SGLT2, FU_finerenone, FU_GLP1,
  elig_RAASi_all, elig_SGLT2i_total, elig_finerenone, elig_GLP1_RA,
  DialysisMonthsSaved_byTreat_SurvCorr,
  DialysisMonthsSaved_byTreatPartial_SurvCorr,
  DialysisMonthsSaved_byFin_SurvCorr
) %>%
    mutate(
        across(
            where(~ all(na.omit(.) %in% c(0, 1))),
            ~ factor(., levels = c(0, 1))
        )
    ) %>%
    tbl_summary(
        type = list(
            everything() ~ "continuous",
            where(is.factor) ~ "categorical"
        ),
        statistic = list(
            all_continuous() ~ "{mean}, [{min}, {max}]",
            all_categorical() ~ "{n} ({p}%)"
        ),
        digits = all_continuous() ~ c(2, 2, 2)
    ) %>%
    modify_caption('*CKD population **eligable for fineranone**. \
Includes treatment used (\"FU_[...]\"), treatment eligibility (\"elig_[...]\"), \
and average [range] dialysis-free months with or without survival correction.*')
```

### GLP1-RA

```{r}
data_ckd %>% dplyr::filter(
  identified_CKD_by_lab == 1, elig_GLP1_RA == 1
) %>% select(
  sex, age,
  FU_RAASi, FU_SGLT2, FU_finerenone, FU_GLP1,
  elig_RAASi_all, elig_SGLT2i_total, elig_finerenone, elig_GLP1_RA,
  DialysisMonthsSaved_byTreat_SurvCorr,
  DialysisMonthsSaved_byTreatPartial_SurvCorr,
  DialysisMonthsSaved_byGLP_SurvCorr
 ) %>%
    mutate(
        across(
            where(~ all(na.omit(.) %in% c(0, 1))),
            ~ factor(., levels = c(0, 1))
        )
    ) %>%
    tbl_summary(
        type = list(
            everything() ~ "continuous",
            where(is.factor) ~ "categorical"
        ),
        statistic = list(
            all_continuous() ~ "{mean}, [{min}, {max}]",
            all_categorical() ~ "{n} ({p}%)"
        ),
        digits = all_continuous() ~ c(2, 2, 2)
    ) %>%
  modify_caption('*CKD population **eligable for GLP1-RA**. \
Includes treatment used (\"FU_[...]\"), treatment eligibility (\"elig_[...]\"), \
and average [range] dialysis-free months with or without survival correction.*')
```

### Sums

```{r}
## all CKD,  all treatment together
data_ckd %>% 
  dplyr::filter(identified_CKD_by_lab == 1) %>% 
  dplyr::summarise(suma = sum(DialysisMonthsSaved_byTreat_SurvCorr))

## all CKD, partial additivity
data_ckd %>% 
  dplyr::filter(identified_CKD_by_lab == 1) %>% 
  dplyr::summarise(suma = sum(DialysisMonthsSaved_byTreatPartial_SurvCorr))

## all CKD,  SGLT treatment
data_ckd %>% 
  dplyr::filter(identified_CKD_by_lab == 1) %>% 
  dplyr::summarise(suma = sum(DialysisMonthsSaved_bySGLT_SurvCorr))

## SGLT eligible,  all treatment
data_ckd %>% 
  dplyr::filter(identified_CKD_by_lab == 1, elig_SGLT2i_total == 1) %>% 
  dplyr::summarise(suma = sum(DialysisMonthsSaved_byTreat_SurvCorr))

## SGLT eligible,  SGLT treatment
data_ckd %>% 
  dplyr::filter(identified_CKD_by_lab == 1, elig_SGLT2i_total == 1) %>% 
  dplyr::summarise(suma = sum(DialysisMonthsSaved_bySGLT_SurvCorr))
```

## Visualisation

### Full additivity

```{r, fig.width=7, fig.height=4}
#| fig.cap: '(A) Proportion of patients with CKD with a positive model-projected gain in dialysis-free time under optimized renoprotective treatment. (B) Distribution of the model-projected gain in dialysis-free time (months) among patients with a positive projected gain'


plotac <- "fig_4"
path <- "gitignore/figures"

tr <- data_ckd %>%
  filter(identified_CKD_by_lab == 1) %>%
  mutate(
    projected_gain = factor(
      DialysisMonthsSaved_byTreat_SurvCorr > 0
    )
  ) %>%
  select(projected_gain)

summary(tr)
mean(tr$projected_gain == "TRUE", na.rm = TRUE)

p1 <- data_ckd %>%
  filter(identified_CKD_by_lab == 1) %>%
  mutate(
    projected_gain = DialysisMonthsSaved_byTreat_SurvCorr > 0
  ) %>%
  ggplot(
    aes(
      x = factor(1),
      fill = projected_gain
    )
  ) +
  geom_bar(position = "fill") +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_fill_manual(
    values = c(
      `FALSE` = "#F4A8A8",
      `TRUE` = "#A8E6A1"
    ),
    labels = c(
      `FALSE` = "Not expected",
      `TRUE` = "Expected"
    )
  ) +
  labs(
    x = "",
    y = "Proportion of CKD patients",
    fill = "Projected gain"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

p2 <- data_ckd %>%
  filter(
    identified_CKD_by_lab == 1,
    DialysisMonthsSaved_byTreat_SurvCorr > 0
  ) %>%
  ggplot(
    aes(
      x = DialysisMonthsSaved_byTreat_SurvCorr
    )
  ) +
  geom_histogram(
    breaks = seq(
      0,
      max(
        data_ckd$DialysisMonthsSaved_byTreat_SurvCorr,
        na.rm = TRUE
      ),
      by = 10
    ),
    fill = "#A8E6A1",
    color = "#4C8A4A"
  ) +
  labs(
    x = "Projected gain in dialysis-free time (months)",
    y = "Number of CKD patients"
  )

fig <- cowplot::plot_grid(
  p1,
  p2,
  rel_widths = c(0.4, 0.6),
  labels = "AUTO"
)

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggplot2::ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 7,
    height = 4
  )
}
```

### Partial additivity

```{r, fig.width=7, fig.height=4}
#| fig.cap: '(A) Proportion of patients with CKD with a positive model-projected gain in dialysis-free time under optimized renoprotective treatment assuming partially additive treatment effects. (B) Distribution of the model-projected gain in dialysis-free time (months) among patients with a positive projected gain'

plotac <- "suppl_fig_8"
path <- "gitignore/figures"

p1 <- data_ckd %>%
  filter(identified_CKD_by_lab == 1) %>%
  mutate(
    projected_gain = DialysisMonthsSaved_byTreatPartial_SurvCorr > 0
  ) %>%
  ggplot(
    aes(
      x = factor(1),
      fill = projected_gain
    )
  ) +
  geom_bar(position = "fill") +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_fill_manual(
    values = c(
      `FALSE` = "#F4A8A8",
      `TRUE` = "#A8E6A1"
    ),
    labels = c(
      `FALSE` = "Not expected",
      `TRUE` = "Expected"
    )
  ) +
  labs(
    x = "",
    y = "Proportion of CKD patients",
    fill = "Projected gain"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

p2 <- data_ckd %>%
  filter(
    identified_CKD_by_lab == 1,
    DialysisMonthsSaved_byTreatPartial_SurvCorr > 0
  ) %>%
  ggplot(
    aes(
      x = DialysisMonthsSaved_byTreatPartial_SurvCorr
    )
  ) +
  geom_histogram(
    breaks = seq(
      0,
      max(
        data_ckd$DialysisMonthsSaved_byTreatPartial_SurvCorr,
        na.rm = TRUE
      ),
      by = 10
    ),
    fill = "#A8E6A1",
    color = "#4C8A4A"
  ) +
  labs(
    x = "Projected gain in dialysis-free time (months)",
    y = "Number of CKD patients"
  )

fig <- cowplot::plot_grid(
  p1,
  p2,
  rel_widths = c(0.4, 0.6),
  labels = "AUTO"
)

fig

if (!dir.exists(path)) {
  dir.create(path, recursive = TRUE)
}

if (!file.exists(file.path(path, paste0(plotac, ".pdf")))) {
  ggplot2::ggsave(
    filename = paste0(plotac, ".pdf"),
    plot = fig,
    path = path,
    device = "pdf",
    width = 7,
    height = 4
  )
}
```


# Reproducibility

```{r}
sessionInfo()
```
